---
title: "Project Part 4"
output: 
  rmdformats::robobook:
    css: "homework.css"
    highlight: pygments
link-citations: yes
---

**Due:** Friday, December 5, 11:59pm on Canvas

**Submission:** A template file is provided, with a rough outline: [project_part_4_template.qmd](https://sta112-f25.github.io/projects/project_part_4_template.qmd)

Once you have completed the report, submit *both* your 
rendered HTML file and the Quarto file to Canvas.

# Overview

Over the course of this semester, you will complete a project in which you use the tools we learn in class to investigate a research question of your choosing. You will then write a report that communicates the results of your analysis and draws a conclusion about your research question. 

In this final part, your task is compile your work from the previous parts of the project into a final written report.

### Overview of project pieces

Initially I had planned to include an oral presentation of your work. However, I have been sufficiently impressed with your project submissions that I don't think that is necessary. I have redistributed some of the points, and made 15% of the project points free.

* Part 1 (20% of project grade): Select a research question and dataset; use EDA to explore data thoroughly
* Part 2 (20% of project grade): Fit a simple linear regression model; report and interpret results in context of research question
* Part 3 (20% of project grade): Add additional explanatory variables to the model; report and interpret results
* Part 4 (25% of project grade): Compile full analyses into final written document, taking into account feedback from earlier parts
* Free 15%

# Your task in Part 4

Based on your work in the previous parts of the project, you will produce a final written report describing your research question, data, analysis, results, and conclusions. You do not need to perform extensive new analysis here; rather, you are combining the pieces from the previous project parts and adding additional context and discussion.

There is no minimum or maximum length required. However, your writing and analysis should be thorough, with your report addressing the research question, and containing all the elements described below.

## Introduction

Write 1-2 paragraphs in which you:

* Motivate the research questions
* Describe where the data came from
* Provide background on the dataset
* Finish by summarizing what you will do in the report

## Data

In the data section, you have three subsections: Data Summary, Univariate EDA, and Multivariate EDA.

### Data Summary

Write a paragraph that describes details of the data:

* What does a row in the data represent? 
* How many rows and columns are there? 
* What information do the variables record? 
* Are there any missing data?

(Make sure to write your summary as a paragraph, not a list, in your report).

### Univariate EDA

Choose **3** variables that are relevant to the research questions. For each variable: make a plot, calculate any appropriate summary statistics, and discuss the distribution in 2--3 sentences. Make sure to add a title and good labels to each plot. Plots and summary statistics depend on the type of variable.

### Multivariate EDA

Using the variables from Univariate EDA, create one or more plots that allow you to address the research questions. You may be able to address both questions with a single plot, or you may need to make a separate plot for each question. Calculate any appropriate summary statistics. Use the plot(s) and summary statistics to address the research questions in a 1--2 paragraphs.

## Modeling

In the modeling section, you have three subsections: Model, Assumption Diagnostics, and Inference.

Here you will use your *final* regression model (chosen after making any adjustments to satisfy regression assumptions, such as variable transformations). Other preliminary or intermediate models which you considered can be included and discussed in the Appendix (below).
  
### Model

First, we need to describe the regression model and give the equation of the fitted model. Using your *final* regression model, write a paragraph in which you:

* Give the population regression model. This should be written using correct notation (remember that $\varepsilon$ is included in the population model, and there are no hats on any variables or parameters). Make sure to use informative variable names (e.g., "Cost" instead of "X").
* Give the equation of the fitted regression model, using correct notation and informative variable names.
* Interpret one or more relevant coefficients of the fitted regression model, and explain what they say about the research question.

### Assumption Diagnostics

Next, we want to verify that the regression assumptions are reasonable for our model. Write 2 - 3 paragraphs in which you:

* Create a residual plot, QQ plot, and studentized residual plot for your fitted model. Make sure to include informative axis labels and captions (with figure numbers) for your figures.
* Use your residual plot and QQ plot to discuss whether the shape, constant variance, and normality assumptions are reasonable. Refer to your figures by their numbers in the text.
* Use your studentized residual plot to identify any potential outliers
* Discuss the independence and randomness assumptions
* Assess multicollinearity for your predictors


### Inference

Finally, we want to use our fitted regression model to address the research question. Write 2 - 3 paragraphs in which you:

* Translate the research question into null and alternative hypotheses about one or more model parameters. Write these hypotheses using correct notation (see "Writing math in R Markdown" below).
* Calculate a test statistics and p-value for your hypotheses.
* Use the p-value to make a conclusion about the research question.
  
## Discussion

In this section, you are providing a conclusion. Write 1–2 paragraphs which summarize what you learned from your analysis, and how it addresses the original research question. Also discuss any limitations to your analysis, and what you other relationships you might explore in future.

## Sources

Provide information on any sources you have used outside the course materials.

* Cite the original source of your data
* Cite any articles you have referenced
* Provide links to any websites you have used for code help
* If generative AI (Chat GPT, Gemini, etc.) was used to help with any part of the assignment, summarize
    * The AI tool used
    * The prompts you used
    * The help provided by the tool
    
## Appendix

If you considered any other models (other than your final model), or created any other exploratory or diagnostic plots that were not included in your main report, you can include these here.


# Appearance and style

### Writing

The report should be written like an article or research paper: in full sentences and paragraphs, with headings for each section. You should not write your report with question numbers or as a list of bullet points. Scientific articles are generally written in third person, though "we" can also be acceptable ("we can see from Figure 1...") in some disciplines.

### Code

As with previous assignments, you will write the document in Quarto. However, in full reports, the only output that should be visible from code chunks are figures and tables. If a code chunk does not produce a figure or table, you can hide it from the knitted document with `include=F`:

````
```{r, include=F}`r ''`

```
````

If a code chunk produces a figure or table, only the figure or table should be visible in the knitted document. You can hide the chunk but display the output with `echo=F, message=F, warning=F`:

````
```{r, echo=F, message=F, warning=F}`r ''`

```
````

### Figures

Figures should have labeled axes, and should be clear and easy to read. Figures should also be captioned and numbered; to caption a figure, use `fig.cap = "..."` in the chunk options. For example (scroll to the right on the code to see it all),

````
```{r, echo=F, message=F, warning=F, fig.cap="Figure 1: Bill depth vs. bill length for penguins near Palmer Station, Antarctica."}`r ''`
penguins |>
  ggplot(aes(x = bill_length_mm, 
             y = bill_depth_mm)) +
  geom_point() +
  labs(x = "Bill length (mm)",
       y = "Bill depth (mm)") +
  theme_bw()
```
````

is displayed as

```{r include=F}
library(tidyverse)
library(palmerpenguins)
```

```{r, echo=F, message=F, warning=F, fig.cap="Figure 1: Bill depth vs. bill length for penguins near Palmer Station, Antarctica."}
penguins |>
  ggplot(aes(x = bill_length_mm, 
             y = bill_depth_mm)) +
  geom_point() +
  labs(x = "Bill length (mm)",
       y = "Bill depth (mm)") +
  theme_bw()
```

Captions should provide enough information to understand what is being plotted, but interpretation can be left to the main text. Refer to figures by their number in the text. Make sure that any figures you include are discussed in the text.

### Writing math in Quarto documents

Doing things like writing down models often involve parameters. Parameters in statistics are generally represented as Greek letters. How can we write Greek letters and other mathematical notation in our Markdown file so they show up when we knit?

If you want to write mathematical notation, we need to tell Markdown, “Hey, we’re going to make a math symbol!” To do that, you use dollar signs. For instance, to make $\widehat{\beta}_1$, you simply put `$\widehathat{\beta}_1$` into the white space (not a chunk) in your Markdown.

Here are some examples of writing math, which you can adapt:

| Math | Code |
| --- | --- |
| $Y = \beta_0 + \beta_1 X + \varepsilon$ | `$Y = \beta_0 + \beta_1 X + \varepsilon$` |
| $\widehat{Y} = -2.2 + 3.7 X$ | `$\widehat{Y} = -2.2 + 3.7 X$` |

