---
title: "Project Part 1"
output: 
  rmdformats::robobook:
    css: "homework.css"
    highlight: pygments
link-citations: yes
---

**Due:** Friday, September 12, 10:00pm on Canvas

**Submission:** A template file is provided, with a rough outline: [project_part_1_template.qmd](https://sta112-f25.github.io/projects/project_part_1_template.qmd)

Once you have completed the report, submit *both* your 
rendered HTML file and the Quarto file to Canvas.

# Overview

Over the course of this semester, you will complete a project in which you use the tools we learn in class to investigate a research question of your choosing. You will then write a report that communicates the results of your analysis and draws a conclusion about your research question, and finally present your work to me. 

This project will be broken into several different parts; I will provide feedback on each part.

In this first part, your task is to identify a research question and a dataset that allows you to address that research question. You will then use exploratory data analysis (EDA) to explore the data and variables. Subsequent parts of the project will involve using linear regression to answer your research question, and then present your results in a written report and an oral interview.

### Overview of project pieces

* Part 1 (15% of project grade): Select a research question and dataset; use EDA to explore data thoroughly
* Part 2 (15% of project grade): Fit a simple linear regression model; report and interpret results in context of research question
* Part 3 (15% of project grade): Add additional explanatory variables to the model; report and interpret results
* Part 4 (25% of project grade): Compile full analyses into final written document, taking into account feedback from earlier parts
* Part 5 (30% of project grade): Oral interview to present your results and processes to Dr. Evans
    * I will be asking questions as you go along the way, just like a client would do who was interested in your results. Be prepared to explain your process and results, and answer questions from me!

# Your task in Part I

Your written report for Part I will contain four sections: Research question, Data Summary, Univariate EDA, and Multivariate EDA. Headings for these sections are included in the template file.

There is no minimum or maximum length required. However, your analysis should be thorough, with your report addressing the research question, and containing all the elements described below.

## Research question and data

Come up with a research question that you can answer with data, and find a dataset that allows you to tackle this research question. 

An example of a research question we have seen in class is from the FiveThirtyEight article on the first day: "Do home teams perform better (score more points compared to the opposition) when there are more fans in the stands?" The dataset in this example is the combination of score data from Basketball Reference, and the fan attendance data.

(You may not use this example, nor any of the other datasets we have worked with in class).

**Requirements:**

* The outcome of interest (e.g., the point differential in the FiveThirtyEight article) must be a *quantitative* variable
* The data must contain at least **5** other variables which you think may be related to the outcome of interest (e.g., fan attendance in the example above), and which help address your research question.
    * At least one of these variables must be categorical
    * At least two of these variables must be quantitative
* Everyone must choose their own data and research question
    * In the unlikely event that two students choose the same data and researc question, the first student to submit the assignment will receive priority

## Data Summary

Write a paragraph that describes details of the data:

* What does a row in the data represent? 
* How many rows and columns are there? 
* What information do the variables record? 
* Are there any missing data?

(Make sure to write your summary as a paragraph, not a list, in your report).

## Univariate EDA

Choose **3** variables that are relevant to the research question. For each variable: make a plot, calculate any appropriate summary statistics, and discuss the distribution in 2--3 sentences. Make sure to add a title and good labels to each plot. Plots and summary statistics depend on the type of variable.

I have summarized how to do univariate EDA below. I have also put together a [data exploration cheat sheet](https://sta112-f25.github.io/homework/data_visualization_summary_cheat_sheet.html) with more information and examples.

**Quantitative variables:** Use a histogram to visualize the distribution of a quantitative variable. Calculate summary statistics (either the mean and standard deviation, or the median and IQR). Describe the distribution based on the plot and summary statistics, making sure to discuss:

* Shape (symmetric vs. skewed, number of modes, location of modes)
* Center (usually mean or median)
* Spread (usually standard deviation or IQR)
* Any unusual features?
* Any potential outliers?

**Categorical variables:** Use a bar chart to visualize the distribution of a categorical variable. When describing the distribution:

* Which category has the most observations? The least?
* Are observations spread relatively evenly across categories, or do one or two categories have the majority of the observations?

(Make sure to write your summary in sentences, not a list, in your report).

## Multivariate EDA

Using the three variables from Univariate EDA, create one or more plots that allow you to address the research question. You may be able to address the question with a single plot, or you may need to make multiple plots. Calculate any appropriate summary statistics. Use the plot(s) and summary statistics to address the research question in a 1--2 paragraphs.

An overview of multivariate EDA is below. See the [data exploration cheat sheet](https://sta112-f25.github.io/homework/data_visualization_summary_cheat_sheet.html) for examples and more information.

**Two quantitative variables:** Visualize the relationship with a scatterplot, and summarize with the correlation if the relationship appears roughly linear. When describing the relationship, reference the scatterplot and correlation, and discuss the following:

* does there appear to be a relationship?
* if so, does the relationship appear to be positive or negative?
* what is the general shape of the relationship? Does it look linear?
* if the relationship looks linear, report the sample correlation coefficient

**One categorical variable, one quantitative variable:** Visualize the relationship with a boxplot or histogram for each level of the categorical variable. Calculate summary statistics for the quantitative variable at each level of the categorical variable (e.g., mean body mass for each penguin species). When describing the relationship, reference the plot and summary statistics, and discuss:

* Is the distribution of the quantitative variable different across levels of the categorical variable? 
* If so, how? (e.g., differences in shape, center, spread)

**Two categorical variables:** Visualize with a stacked bar chart, and describe the plot:

* Which combination of categories has the most observations? The least?
* Are there any combinations which do not appear in the data?
* Is the distribution for the second variable the same for each level of the first variable?

**Two quantitative variables, one categorical:** Visualize with a scatterplot, and color and shape the points according to the categorical variable. Calculate correlation between the quantitative variables for each level of the categorical variable. Describe the plot and summary statistics:

* Does the relationship between the two quantitative variables differ for different levels of the categorical variable? If so, how?

# A few ideas for data sources

Here are some ideas for data sources, but you are also welcome to find data elsewhere!

* [CORGIS: The Collection of Really Great, Interesting, Situated Dataset](https://corgis-edu.github.io/corgis/csv/)
* [Awesome Public Datasets](https://github.com/awesomedata/awesome-public-datasets) - large but categorized list by topic and data type
* [Driven Data](https://www.drivendata.org/competitions/)
* You can create data from the US Census Bureau using [tidyCensus](https://walker-data.com/tidycensus/articles/basic-usage.html)
* [Data on Fires in the US and resultant damage](https://www.usfa.fema.gov/statistics/data-sets/)
* [Data Is Plural](https://www.data-is-plural.com/) -- Weekly feed (with archive) of interesting datasets
* [AWS Data Exchange - Open Data](https://aws.amazon.com/marketplace/search/results?CONTRACT_TYPE=OPEN_DATA_LICENSES&filters=CONTRACT_TYPE)
* The OpenIntro textbooks have a great [list of datasets](https://www.openintro.org/data/)
* Shonda Kuiper has compiled a [list](https://stat2labs.sites.grinnell.edu/DataResources.html) of different data sources
* The [TidyTuesday](https://github.com/rfordatascience/tidytuesday) project has a long list of many different datasets
* The [nhanesA](https://cran.r-project.org/web/packages/nhanesA/vignettes/Introducing_nhanesA.html) package allows you to explore and retrieve data from the National Health and Nutrition Examination Survey
* [College ScoreCard data](https://collegescorecard.ed.gov/) for many different institutions
* [American time use survey data](https://github.com/swang87/atus)
* [Lahman baseball data](https://cran.r-project.org/web/packages/Lahman/index.html)
* CDC [Behaviorial Risk Factor Surveillance System](https://www.cdc.gov/brfss/) data
* National Park [visitor use statistics](https://irma.nps.gov/Stats/)


# Appearance and style

### Writing

The report should be written like an article or research paper: in full sentences and paragraphs, with headings for each section. You should not write your report with question numbers or as a list of bullet points. Scientific articles are generally written in third person, though "we" can also be acceptable ("we can see from Figure 1...") in some disciplines.

### Code

As with previous assignments, you will write the document in Quarto. However, in full reports, the only output that should be visible from code chunks are figures and tables. If a code chunk does not produce a figure or table, you can hide it from the knitted document with `include=F`:

````
```{r, include=F}`r ''`

```
````

If a code chunk produces a figure or table, only the figure or table should be visible in the knitted document. You can hide the chunk but display the output with `echo=F, message=F, warning=F`:

````
```{r, echo=F, message=F, warning=F}`r ''`

```
````

### Figures

Figures should have labeled axes, and should be clear and easy to read. Figures should also be captioned and numbered; to caption a figure, use `fig.cap = "..."` in the chunk options. For example (scroll to the right on the code to see it all),

````
```{r, echo=F, message=F, warning=F, fig.cap="Figure 1: Bill depth vs. bill length for penguins near Palmer Station, Antarctica."}`r ''`
penguins |>
  ggplot(aes(x = bill_length_mm, 
             y = bill_depth_mm)) +
  geom_point() +
  labs(x = "Bill length (mm)",
       y = "Bill depth (mm)") +
  theme_bw()
```
````

is displayed as

```{r include=F}
library(tidyverse)
library(palmerpenguins)
```

```{r, echo=F, message=F, warning=F, fig.cap="Figure 1: Bill depth vs. bill length for penguins near Palmer Station, Antarctica."}
penguins |>
  ggplot(aes(x = bill_length_mm, 
             y = bill_depth_mm)) +
  geom_point() +
  labs(x = "Bill length (mm)",
       y = "Bill depth (mm)") +
  theme_bw()
```

Captions should provide enough information to understand what is being plotted, but interpretation can be left to the main text. Refer to figures by their number in the text. Make sure that any figures you include are discussed in the text.


# Checklist

* Change name in Quarto YAML
* Four sections: Research question, Data Summary, Univariate EDA, Multivariate EDA
* Research question: a description of the research question you will investigate
  * Chosen data must have quantitative outcome of interest
  * Chosen data must have at least 5 other variables, including at least one categorical variable and two quantitative variables
* Data Summary:
  * What does a row in the data represent? 
  * How many rows and columns are there? 
  * What variables are in the data?
  * Are there any missing data?
* Univariate EDA:
  * Pick three variables *relevant to the research question*
  * Plot, appropriate summary statistics, and describe distribution for each variable
  * See above for details on appropriate plots and summary statistics, and what to discuss when describing a distribution
* Multivariate EDA:
  * One or more plots that examine the relationships between the three variables from univariate EDA
  * Plots allow you to address the research question
  * Discuss plots in context of research question; see above for more details
* Writing:
  * Each section written as a paragraph, with full sentences
  * No bullet points or lists
* Code:
  * Code is included in the Quarto document, but *not* displayed in the rendered HTML
  * The only R output shown in the report is figures and tables
* Figures:
  * Include meaningful caption and axis labels (see above for details on including a caption)
  * Each figure included in the report is discussed in your writing, and referred to by its figure number
* Submission:
  * Submit *both* the rendered HMTL and the .qmd file to Canvas
  * Deadline is Friday, September 12 at 10pm