---
title: "Project Part 2"
output: 
  rmdformats::robobook:
    css: "homework.css"
    highlight: pygments
link-citations: yes
---

**Due:** Friday, October 3, 10:00pm on Canvas

**Submission:** A template file is provided, with a rough outline: [project_part_2_template.qmd](https://sta112-f25.github.io/projects/project_part_2_template.qmd)

Once you have completed the report, submit *both* your 
rendered HTML file and the Quarto file to Canvas.

# Overview

Over the course of this semester, you will complete a project in which you use the tools we learn in class to investigate a research question of your choosing. You will then write a report that communicates the results of your analysis and draws a conclusion about your research question, and finally present your work to me. 

This project will be broken into several different parts; I will provide feedback on each part.

In this second part, your task is to begin to address your research question by fitting an appropriate linear regression model.

### Overview of project pieces

* Part 1 (15% of project grade): Select a research question and dataset; use EDA to explore data thoroughly
* Part 2 (15% of project grade): Fit a simple linear regression model; report and interpret results in context of research question
* Part 3 (15% of project grade): Add additional explanatory variables to the model; report and interpret results
* Part 4 (25% of project grade): Compile full analyses into final written document, taking into account feedback from earlier parts
* Part 5 (30% of project grade): Oral interview to present your results and processes to Dr. Evans
    * I will be asking questions as you go along the way, just like a client would do who was interested in your results. Be prepared to explain your process and results, and answer questions from me!

# Your task in Part II

Based on your exploratory data analysis from Part 1, fit a linear regression model to address your research question. Then, use a residual plot to assess the shape and constant variance assumptions, and a studentized residual plot to check for potential outliers. Make any additional adjustments to your model based on your diagnostic plots. Finally, interpret the fitted model: discuss the estimated slope in context of the research question, and report and interpret the model's RMSE.

There is no minimum or maximum length required. However, your analysis should be thorough, with your report addressing the research question, and containing all the elements described below.

## Model

First, we need to describe the regression model and give the equation of the fitted line. Write a short paragraph in which you:

* Give the population regression model you will use to address your research question. This should be written using correct notation (remember that $\varepsilon$ is included in the population model, and there are no hats on any variables or parameters). Make sure to use informative variable names. See "Writing math in Quarto" below.
* Give the equation of the fitted regression line, using correct notation and informative variable names (remember to provide the actual values of $\widehat{\beta}_0$ and $\widehat{\beta}_1$).

## Assumption Diagnostics

Next, we want to assess whether the regression assumptions are reasonable for our model. Write 1 - 2 paragraphs in which you:

* Create a residual plot and studentized residual plot for your fitted model. Make sure to include informative axis labels and captions (with figure numbers) for your figures.
* Use your residual plot to discuss whether the shape and constant variance assumptions are reasonable. Refer to your figures by their numbers in the text.
* Use your studentized residual plot to identify any potential outliers
* Make any changes to your regression model based on the residual plot; if you have made changes, report the equation of your final fitted model

## Interpretation

Write a short paragraph in which you:

* Interpret the estimated slope in your final fitted model, in context of the data and research question
* Discuss the RMSE of the fitted model

## Sources

Provide information on any sources you have used outside the course materials.

* Cite any articles you have referenced
* Provide links to any websites you have used for code help
* If generative AI (Chat GPT, Gemini, etc.) was used to help with any part of the assignment, summarize
    * The AI tool used
    * The prompts you used
    * The help provided by the tool


# Appearance and style

### Creating a studentized residual plot

Here is example code for creating a studentized residual plot, using the Florida voting data as an example:

```{r, message=F, warning=F}
library(tidyverse)
library(Stat2Data)
data("PalmBeach")

vote_lm <- lm(Buchanan ~ Bush, data = PalmBeach)

data.frame(studentized_resids = rstudent(vote_lm),
         yhat = predict(vote_lm)) |>
  ggplot(aes(x = yhat, y = studentized_resids)) +
  geom_point() +
  geom_abline(slope = 0, intercept = 0, 
              color = "blue", lwd=1.2) +
  geom_abline(slope = 0, intercept = c(-3, -2, 2, 3),
              color = "orange", lty = 2, lwd=1.2) +
  labs(x = "Predicted votes for Buchanan",
       y = "Studentized residuals") +
  theme_bw()
```


### Writing

The report should be written like an article or research paper: in full sentences and paragraphs, with headings for each section. You should not write your report with question numbers or as a list of bullet points. Scientific articles are generally written in third person, though "we" can also be acceptable ("we can see from Figure 1...") in some disciplines.

### Code

As with previous assignments, you will write the document in Quarto. However, in full reports, the only output that should be visible from code chunks are figures and tables. If a code chunk does not produce a figure or table, you can hide it from the knitted document with `include=F`:

````
```{r, include=F}`r ''`

```
````

If a code chunk produces a figure or table, only the figure or table should be visible in the knitted document. You can hide the chunk but display the output with `echo=F, message=F, warning=F`:

````
```{r, echo=F, message=F, warning=F}`r ''`

```
````

### Figures

Figures should have labeled axes, and should be clear and easy to read. Figures should also be captioned and numbered; to caption a figure, use `fig.cap = "..."` in the chunk options. For example (scroll to the right on the code to see it all),

````
```{r, echo=F, message=F, warning=F, fig.cap="Figure 1: Bill depth vs. bill length for penguins near Palmer Station, Antarctica."}`r ''`
penguins |>
  ggplot(aes(x = bill_length_mm, 
             y = bill_depth_mm)) +
  geom_point() +
  labs(x = "Bill length (mm)",
       y = "Bill depth (mm)") +
  theme_bw()
```
````

is displayed as

```{r include=F}
library(tidyverse)
library(palmerpenguins)
```

```{r, echo=F, message=F, warning=F, fig.cap="Figure 1: Bill depth vs. bill length for penguins near Palmer Station, Antarctica."}
penguins |>
  ggplot(aes(x = bill_length_mm, 
             y = bill_depth_mm)) +
  geom_point() +
  labs(x = "Bill length (mm)",
       y = "Bill depth (mm)") +
  theme_bw()
```

Captions should provide enough information to understand what is being plotted, but interpretation can be left to the main text. Refer to figures by their number in the text. Make sure that any figures you include are discussed in the text.

### Writing math in Quarto documents

Doing things like writing down models often involve parameters. Parameters in statistics are generally represented as Greek letters. How can we write Greek letters and other mathematical notation in our Markdown file so they show up when we knit?

If you want to write mathematical notation, we need to tell Markdown, “Hey, we’re going to make a math symbol!” To do that, you use dollar signs. For instance, to make $\widehat{\beta}_1$, you simply put `$\widehathat{\beta}_1$` into the white space (not a chunk) in your Markdown.

Here are some examples of writing math, which you can adapt:

| Math | Code |
| --- | --- |
| $Y = \beta_0 + \beta_1 X + \varepsilon$ | `$Y = \beta_0 + \beta_1 X + \varepsilon$` |
| $\widehat{Y} = -2.2 + 3.7 X$ | `$\widehat{Y} = -2.2 + 3.7 X$` |

# Checklist

* Change name in Quarto YAML
* Four sections: Model, Diagnostics, Interpretation, and Sources
* Model
  * Report the population regression model
  * Report the equation of the fitted regression model
* Assumption Diagnostics
  * Residual plot and studentized residual plot
  * Discuss shape and constant variance
  * Identify any potential outliers
  * If any changes are made to model report equation of final fitted model
* Interpretation
  * Discuss and interpret fitted model in context of research question
  * Discuss RMSE of fitted model
* Sources:
  * Cite any articles
  * Cite any websites which provided code or statistical help
  * Cite any AI use (tool, prompts, help provided)
* Writing:
  * Each section written as a paragraph, with full sentences
  * No bullet points or lists
  * Math equations are written nicely, with correct notation (see "Writing math in Quarto")
* Code:
  * Code is included in the Quarto document, but *not* displayed in the rendered HTML
  * The only R output shown in the report is figures and tables
* Figures:
  * Include meaningful caption and axis labels (see above for details on including a caption)
  * Each figure included in the report is discussed in your writing, and referred to by its figure number
* Submission:
  * Submit *both* the rendered HMTL and the .qmd file to Canvas
  * Deadline is Friday, October 3 at 10pm