---
title: Introduction to multiple linear regression
output:
  xaringan::moon_reader:
    css: "lab-slides.css"
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
editor_options: 
  chunk_output_type: console
---

### Benefits of multiple linear regression

.large[
.question[
What are some benefits of including multiple predictors in my regression model?
]
]

---

### Benefits of multiple linear regression

.large[
.question[
What are some benefits of including multiple predictors in my regression model?
]
]

.large[
Two main reasons:

* Improve the ability to predict the response
* Account for other variables that may change the relationship (e.g., penguin species)
]

---

### Improving model fit

```{r echo=F, message=F, warning=F, fig.width=7, fig.height=4.5, fig.align='center'}

library(palmerpenguins)
library(tidyverse)
library(latex2exp)
library(Stat2Data)
library(knitr)

penguins <- penguins |>
  drop_na(bill_length_mm, species, body_mass_g)

penguins %>%
  ggplot(aes(x = body_mass_g, y = bill_length_mm)) +
  geom_point(size=2) +
  geom_smooth(se=F, method="lm", lwd=1.5) +
  labs(x = "Body mass (g)",
       y = "Bill length (mm)") +
  theme_bw() +
  theme(text = element_text(size = 25))
```

.large[
$$\widehat{\text{bill length}} = 27.15 + 0.004 \ \text{body mass}$$

.question[
How do I measure how well body mass predicts bill length?
]
]

---

### Improving model fit

```{r echo=F, message=F, warning=F, fig.width=7, fig.height=4.5, fig.align='center'}
penguins %>%
  ggplot(aes(x = body_mass_g, y = bill_length_mm)) +
  geom_point(size=2) +
  geom_smooth(se=F, method="lm", lwd=1.5) +
  labs(x = "Body mass (g)",
       y = "Bill length (mm)") +
  theme_bw() +
  theme(text = element_text(size = 25))
```

.large[
$$\widehat{\text{bill length}} = 27.15 + 0.004 \ \text{body mass}$$

$R^2 = 0.348$, $R^2_{adj} = 0.346$
]

---

### Improving model fit

```{r echo=F, message=F, warning=F, fig.width=7, fig.height=4.5, fig.align='center'}
length_lm <- lm(bill_length_mm ~ species + body_mass_g, 
                data = penguins)

penguins %>%
  mutate(yhat = predict(length_lm)) %>%
  ggplot(aes(x = body_mass_g,
             y = bill_length_mm,
             color = species)) +
  geom_point(size = 2, alpha = 0.5) +
  geom_line(aes(y = yhat), 
            lwd = 1.5) +
  labs(x = "Body mass (g)",
       y = "Bill length (mm)",
       color = "Species") +
  theme_bw() +
  theme(text = element_text(size = 25))
```

.large[
But we should probably fit a line for each species...

.question[
Do you think $R^2_{adj}$ will increase or decrease when we add species to the model?
]
]

---

### Improving model fit

```{r echo=F, message=F, warning=F, fig.width=7, fig.height=4.5, fig.align='center'}
length_lm <- lm(bill_length_mm ~ species + body_mass_g, 
                data = penguins)

penguins %>%
  mutate(yhat = predict(length_lm)) %>%
  ggplot(aes(x = body_mass_g,
             y = bill_length_mm,
             color = species)) +
  geom_point(size = 2, alpha = 0.5) +
  geom_line(aes(y = yhat), 
            lwd = 1.5) +
  labs(x = "Body mass (g)",
       y = "Bill length (mm)",
       color = "Species") +
  theme_bw() +
  theme(text = element_text(size = 25))
```

.large[
\begin{align}
\widehat{\text{bill length}} &= 24.91 + 9.91 \ \text{IsChinstrap} + 3.54 \ \text{IsGentoo} \\
& \hspace{0.5cm} + 0.0038 \ \text{body mass}
\end{align}
]

.large[
$R^2 = 0.806$, $R^2_{adj} = 0.804$
]

---

### Improving model fit

.pull-left[
```{r echo=F, message=F, warning=F, fig.width=6, fig.height=4, fig.align='center'}
penguins %>%
  ggplot(aes(x = body_mass_g, y = bill_length_mm)) +
  geom_point(size=2) +
  geom_smooth(se=F, method="lm", lwd=1.5) +
  labs(x = "Body mass (g)",
       y = "Bill length (mm)") +
  theme_bw() +
  theme(text = element_text(size = 25))
```

.large[
$R^2_{adj} = 0.346$
]
]

.pull-right[
```{r echo=F, message=F, warning=F, fig.width=6, fig.height=4, fig.align='center'}

penguins %>%
  mutate(yhat = predict(length_lm)) %>%
  ggplot(aes(x = body_mass_g,
             y = bill_length_mm,
             color = species)) +
  geom_point(size = 2, alpha = 0.5) +
  geom_line(aes(y = yhat), 
            lwd = 1.5) +
  labs(x = "Body mass (g)",
       y = "Bill length (mm)",
       color = "Species") +
  theme_bw() +
  theme(text = element_text(size = 25))
```

.large[
$R^2_{adj} = 0.804$
]
]

.large[
* Our ability to predict bill length increases substantially when we add Species to the model
* Make sure to compare $R^2_{adj}$ (rather than $R^2$ ) when comparing models
]

---

### Accounting for other variables

.pull-left[
```{r echo=F, message=F, warning=F, fig.width=6, fig.height=4, fig.align='center'}
penguins %>%
  ggplot(aes(x = body_mass_g, y = bill_length_mm)) +
  geom_point(size=2) +
  geom_smooth(se=F, method="lm", lwd=1.5) +
  labs(x = "Body mass (g)",
       y = "Bill length (mm)") +
  theme_bw() +
  theme(text = element_text(size = 25))
```

]

.pull-right[
```{r echo=F, message=F, warning=F, fig.width=6, fig.height=4, fig.align='center'}

penguins %>%
  mutate(yhat = predict(length_lm)) %>%
  ggplot(aes(x = body_mass_g,
             y = bill_length_mm,
             color = species)) +
  geom_point(size = 2, alpha = 0.5) +
  geom_line(aes(y = yhat), 
            lwd = 1.5) +
  labs(x = "Body mass (g)",
       y = "Bill length (mm)",
       color = "Species") +
  theme_bw() +
  theme(text = element_text(size = 25))
```

]

.large[
The slope looks pretty similar regardless of whether we add Species to the model. But that isn't always the case...
]

---

### Accounting for other variables

```{r echo=F, message=F, warning=F, fig.width=7, fig.height=5, fig.align='center'}
penguins_raw_no_nas <- penguins_raw %>%
  mutate(species_short = ifelse(Species == "Adelie Penguin (Pygoscelis adeliae)", "Adelie",
                                ifelse(Species == "Gentoo penguin (Pygoscelis papua)", "Gentoo", "Chinstrap"))) %>%
  select(`Culmen Length (mm)`, `Culmen Depth (mm)`, species_short) %>%
  drop_na()

penguins_raw_no_nas %>%
  ggplot(aes(x = `Culmen Length (mm)`, y = `Culmen Depth (mm)`)) +
  geom_point(size=2) +
  geom_smooth(se=F, method="lm", lwd=1.5) +
  theme_bw() +
  theme(text = element_text(size=25))
```

.large[
There appears to be a weak, negative relationship between culmen length and culmen depth.
]

---

### Accounting for other variables

```{r echo=F, message=F, warning=F, fig.width=7, fig.height=5, fig.align='center'}
penguins_raw_no_nas %>%
  ggplot(aes(x = `Culmen Length (mm)`, y = `Culmen Depth (mm)`,
             color = species_short)) +
  geom_point(size=2, alpha = 0.5) +
  geom_smooth(se=F, method="lm", lwd=1.5) +
  labs(color = "Species") +
  theme_bw() +
  theme(text = element_text(size=25))
```

.large[
But when we include Species in the model, we see *positive* relationships between culmen length and culmen depth.
]

---

### Simpson's paradox

.pull-left[
```{r echo=F, message=F, warning=F, fig.width=6, fig.height=4, fig.align='center'}

penguins_raw_no_nas %>%
  ggplot(aes(x = `Culmen Length (mm)`, y = `Culmen Depth (mm)`)) +
  geom_point(size=2) +
  geom_smooth(se=F, method="lm", lwd=1.5) +
  theme_bw() +
  theme(text = element_text(size=25))
```
]

.pull-right[
```{r echo=F, message=F, warning=F, fig.width=6, fig.height=4, fig.align='center'}
penguins_raw_no_nas %>%
  ggplot(aes(x = `Culmen Length (mm)`, y = `Culmen Depth (mm)`,
             color = species_short)) +
  geom_point(size=2, alpha = 0.5) +
  geom_smooth(se=F, method="lm", lwd=1.5) +
  labs(color = "Species") +
  theme_bw() +
  theme(text = element_text(size=25))
```
]

.large[
**Simpson's paradox:** The apparent relationship between variables changes when you look within groups
]

---

### Accounting for other variables

.pull-left[
```{r echo=F, message=F, warning=F, fig.width=6, fig.height=4, fig.align='center'}

penguins_raw_no_nas %>%
  ggplot(aes(x = `Culmen Length (mm)`, y = `Culmen Depth (mm)`)) +
  geom_point(size=2) +
  geom_smooth(se=F, method="lm", lwd=1.5) +
  theme_bw() +
  theme(text = element_text(size=25))
```
]

.pull-right[
```{r echo=F, message=F, warning=F, fig.width=6, fig.height=4, fig.align='center'}
penguins_raw_no_nas %>%
  ggplot(aes(x = `Culmen Length (mm)`, y = `Culmen Depth (mm)`,
             color = species_short)) +
  geom_point(size=2, alpha = 0.5) +
  geom_smooth(se=F, method="lm", lwd=1.5) +
  labs(color = "Species") +
  theme_bw() +
  theme(text = element_text(size=25))
```
]

.large[
* The model *without* Species estimates the relationship between culmen length and culmen depth in the population of all Adelie, Chinstrap, and Gentoo penguins
* The model *with* Species estimates the relationship between culmen length and culmen depth *after accounting for species*
]