---
title: Regression with a categorical predictor
output:
  xaringan::moon_reader:
    css: "lab-slides.css"
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
editor_options: 
  chunk_output_type: console
---

### Reminders and agenda

.large[
* Exam 1 next Friday (10/10)
    * Covers material through the last class (9/29)
* Exam 1 review day next Wednesday (10/8)
* Studying advice:
    * Practice questions on the course website
    * Review previous class activities
    * Review previous homework assignments
* Today: thinking about categorical explanatory variables
]

---

### Last time: Class activity

.large[
Data on doctors and hospitals from a random sample of 53 US counties:

* `County`: the name and state of the county
* `MDs`: the number of medical doctors in the county
* `Hospitals`: number of hospitals in the county
* `Beds`: the number of beds in the hospital
]

---

### Class activity

```{r, echo=F, message=F, fig.width=7, fig.height=5, fig.align='center'}
library(tidyverse)
library(Stat2Data)
data("CountyHealth")
health_lm <- lm(MDs ~ Hospitals, data = CountyHealth)
CountyHealth %>%
  mutate(yhat = predict(health_lm),
         residuals = residuals(health_lm)) %>%
  ggplot(aes(x = yhat, y = residuals)) +
  geom_point(size=2) +
  geom_abline(slope = 0, intercept = 0, 
              color = "blue", lwd = 1.2) +
  labs(x = "Predicted number of doctors",
       y = "Residual") +
  theme_bw() +
  theme(text = element_text(size = 25))
```

.large[
.question[
Do the shape and constant variance assumptions look reasonable?
]
]

---

### Class activity

```{r, echo=F, message=F, fig.width=7, fig.height=5, fig.align='center'}


CountyHealth %>%
  mutate(yhat = predict(health_lm),
         residuals = residuals(health_lm)) %>%
  ggplot(aes(x = yhat, y = residuals)) +
  geom_point(size=2) +
  geom_abline(slope = 0, intercept = 0, 
              color = "blue", lwd = 1.2) +
  labs(x = "Predicted number of doctors",
       y = "Residual") +
  theme_bw() +
  theme(text = element_text(size = 25))
```

.large[
* Shape may be slightly violated
* Constant variance is definitely violated
]

---

### Class activity

```{r, echo=F, message=F, fig.width=7, fig.height=5, fig.align='center'}
CountyHealth %>%
  mutate(residuals = residuals(health_lm)) %>%
  ggplot(aes(sample = residuals)) +
  geom_qq(size=2) +
  geom_qq_line() +
  labs(x = "Theoretical normal quantiles",
       y = "Observed residual quantiles") +
  theme_classic() +
  theme(text = element_text(size=25))
```

.large[
.question[
Does the normality assumption look reasonable?
]
]

---

### Class activity

```{r, echo=F, message=F, fig.width=7, fig.height=5, fig.align='center'}
CountyHealth %>%
  mutate(residuals = residuals(health_lm)) %>%
  ggplot(aes(sample = residuals)) +
  geom_qq(size=2) +
  geom_qq_line() +
  labs(x = "Theoretical normal quantiles",
       y = "Observed residual quantiles") +
  theme_classic() +
  theme(text = element_text(size=25))
```

.large[
The normality assumption is definitely violated.
]

---

### Transformation

.large[
Trying a square root transformation of the response:
]

```{r, echo=F, message=F, fig.width=7, fig.height=5}
health_lm <- lm(sqrt(MDs) ~ Hospitals, data = CountyHealth)
CountyHealth %>%
  mutate(yhat = predict(health_lm),
         residuals = residuals(health_lm)) %>%
  ggplot(aes(x = yhat, y = residuals)) +
  geom_point(size=2) +
  geom_abline(slope = 0, intercept = 0, 
              color = "blue", lwd = 1.2) +
  labs(x = "Predicted sqrt(number of doctors)",
       y = "Residual") +
  theme_bw() +
  theme(text = element_text(size = 20))
```

.large[
.question[
How do the shape and constant variance assumptions look now?
]
]

---

### Transformation

.large[
QQ plot with the transformed data:
]

```{r, echo=F, message=F, fig.width=7, fig.height=5, fig.align='center'}
CountyHealth %>%
  mutate(residuals = residuals(health_lm)) %>%
  ggplot(aes(sample = residuals)) +
  geom_qq(size=2) +
  geom_qq_line() +
  labs(x = "Theoretical normal quantiles",
       y = "Observed residual quantiles") +
  theme_classic() +
  theme(text = element_text(size=20))
```

.large[
.question[
How does the normality assumption look now?
]
]


---

### Class activity

.large[
.question[
Do the independence and randomness assumptions seem reasonable for the `CountyHealth` data?
]
]

---

### Class activity

.large[
.question[
Do the independence and randomness assumptions seem reasonable for the `CountyHealth` data?
]
]

.large[
Yes, both seem reasonable.

* Independence: reasonable to assume the number of hospitals/doctors for one county doesn't affect the number of hospitals/doctors for another county
* Randomness: we are told the data come from a random sample
]

---

### Regression so far

.large[
Simple linear regression:

$$y = \beta_0 + \beta_1 x + \varepsilon$$
Polynomial regression:

$$y = \beta_0 + \beta_1 x + \beta_2 x^2 + \cdots + \beta_k x^k + \varepsilon$$

where $x$ and $y$ are both **quantitative**.

.question[
What if $x$ is *categorical*?
]
]

---

### Categorical predictor

```{r echo=F, message=F, warning=F, fig.align='center', fig.width=6, fig.height=4}
library(palmerpenguins)
library(tidyverse)

penguins_no_nas <- penguins %>%
  drop_na()

penguins_no_nas %>%
  ggplot(aes(x = species, y = bill_length_mm)) +
  geom_boxplot(lwd=1.5) +
  theme_classic() +
  labs(x = "Species", 
       y = "Bill length (mm)") +
  theme(text = element_text(size = 25))
```

.large[
* How do we model the relationship between species and bill length?
* Can we test for a relationship between bill length and species?
]

---

### Categorical predictor

.large[
Imagine population like this:
]

```{r echo=F, message=F, warning=F, fig.align='center', fig.width=6, fig.height=4}

set.seed(1)
penguins_no_nas %>%
  ggplot(aes(x = species, y = bill_length_mm)) +
  geom_jitter(alpha = 0.5, size=2) +
  theme_classic() +
  geom_segment(aes(x = 0.6, y = 38.8, xend = 1.4, yend = 38.8), lwd=1.5, color="blue") + 
  geom_segment(aes(x = 1.6, y = 48.8, xend = 2.4, yend = 48.8), lwd=1.5, color="blue") + 
  geom_segment(aes(x = 2.6, y = 47.6, xend = 3.4, yend = 47.6), lwd=1.5, color="blue") + 
  labs(x = "Species", 
       y = "Bill length (mm)") +
  theme(text = element_text(size = 25))
```

.large[
If $\text{species} = \text{Adelie}$, 

$$\hspace{1cm} \text{bill length} = \text{mean Adelie bill length} + \varepsilon$$
]

---

### Categorical predictor

.large[
Imagine population like this:
]

```{r echo=F, message=F, warning=F, fig.align='center', fig.width=6, fig.height=4}

set.seed(1)
penguins_no_nas %>%
  ggplot(aes(x = species, y = bill_length_mm)) +
  geom_jitter(alpha = 0.5, size=2) +
  theme_classic() +
  geom_segment(aes(x = 0.6, y = 38.8, xend = 1.4, yend = 38.8), lwd=1.5, color="blue") + 
  geom_segment(aes(x = 1.6, y = 48.8, xend = 2.4, yend = 48.8), lwd=1.5, color="blue") + 
  geom_segment(aes(x = 2.6, y = 47.6, xend = 3.4, yend = 47.6), lwd=1.5, color="blue") + 
  labs(x = "Species", 
       y = "Bill length (mm)") +
  theme(text = element_text(size = 25))
```

.large[
If $\text{species} = \text{Chinstrap}$, 

$$\hspace{0.3cm} \text{bill length} = \text{mean Chinstrap bill length} + \varepsilon$$
]

---

### Categorical predictor

.large[
Imagine population like this:
]

```{r echo=F, message=F, warning=F, fig.align='center', fig.width=6, fig.height=4}

set.seed(1)
penguins_no_nas %>%
  ggplot(aes(x = species, y = bill_length_mm)) +
  geom_jitter(alpha = 0.5, size=2) +
  theme_classic() +
  geom_segment(aes(x = 0.6, y = 38.8, xend = 1.4, yend = 38.8), lwd=1.5, color="blue") + 
  geom_segment(aes(x = 1.6, y = 48.8, xend = 2.4, yend = 48.8), lwd=1.5, color="blue") + 
  geom_segment(aes(x = 2.6, y = 47.6, xend = 3.4, yend = 47.6), lwd=1.5, color="blue") + 
  labs(x = "Species", 
       y = "Bill length (mm)") +
  theme(text = element_text(size = 25))
```

.large[

If $\text{species} = \text{Gentoo}$, 

$$\hspace{1cm} \text{bill length} = \text{mean Gentoo bill length} + \varepsilon$$
]


---

### Categorical predictor

```{r echo=F, message=F, warning=F, fig.align='center', fig.width=5, fig.height=3}

library(latex2exp)

set.seed(1)
penguins_no_nas %>%
  ggplot(aes(x = species, y = bill_length_mm)) +
  geom_jitter(alpha = 0) +
  theme_classic() +
  geom_segment(aes(x = 0.6, y = 38, xend = 1.4, yend = 38), lwd=1.5, color="blue") + 
  geom_segment(aes(x = 1.6, y = 50, xend = 2.4, yend = 50), lwd=1.5, color="blue") + 
  geom_segment(aes(x = 2.6, y = 47, xend = 3.4, yend = 47), lwd=1.5, color="blue") + 
  labs(x = "Species", 
       y = "Bill length (mm)") +
  theme(text = element_text(size = 25)) +
  theme(axis.title.y = element_blank()) +
  scale_y_continuous(breaks = c(38),
                     labels = c(TeX("$\\beta_0")))
```

.large[
If $\text{species} = \text{Adelie}, \hspace{1cm} \text{bill length} = \beta_0 + \varepsilon$

* $\beta_0$ = true mean bill length of Adelie penguins (in the population)
]

---

### Categorical predictor

```{r echo=F, message=F, warning=F, fig.align='center', fig.width=5, fig.height=3}

library(latex2exp)

set.seed(1)
penguins_no_nas %>%
  ggplot(aes(x = species, y = bill_length_mm)) +
  geom_jitter(alpha = 0) +
  theme_classic() +
  geom_segment(aes(x = 0.6, y = 38, xend = 1.4, yend = 38), lwd=1.5, color="blue") + 
  geom_segment(aes(x = 1.6, y = 50, xend = 2.4, yend = 50), lwd=1.5, color="blue") + 
  geom_segment(aes(x = 2.6, y = 47, xend = 3.4, yend = 47), lwd=1.5, color="blue") + 
  labs(x = "Species", 
       y = "Bill length (mm)") +
  theme(text = element_text(size = 25)) +
  theme(axis.title.y = element_blank()) +
  scale_y_continuous(breaks = c(38),
                     labels = c(TeX("$\\beta_0")))
```

.large[
If $\text{species} = \text{Adelie}, \hspace{1cm} \text{bill length} = \beta_0 + \varepsilon$

* $\beta_0$ = true mean bill length of Adelie penguins (in the population)

.question[
What notation could I use for Chinstrap penguins?
]
]

---

### Categorical predictor

```{r echo=F, message=F, warning=F, fig.align='center', fig.width=5, fig.height=3}

set.seed(1)
penguins_no_nas %>%
  ggplot(aes(x = species, y = bill_length_mm)) +
  geom_jitter(alpha = 0) +
  theme_classic() +
  geom_segment(aes(x = 0.6, y = 38, xend = 1.4, yend = 38), lwd=1.5, color="blue") + 
  geom_segment(aes(x = 1.6, y = 50, xend = 2.4, yend = 50), lwd=1.5, color="blue") + 
  geom_segment(aes(x = 2.6, y = 47, xend = 3.4, yend = 47), lwd=1.5, color="blue") + 
  labs(x = "Species", 
       y = "Bill length (mm)") +
  theme(text = element_text(size = 25)) +
  theme(axis.title.y = element_blank()) +
  scale_y_continuous(breaks = c(38, 50),
                     labels = c(TeX("$\\beta_0"), 
                                TeX("$\\beta_0 + \\beta_1"))) +
  geom_segment(aes(x = 1.45, y = 38,
                   xend = 1.45, yend = 50),
               arrow = arrow(length = unit(0.3, "cm"))) +
  annotate("text", x = 1.3, y = 44, 
           label = TeX("$\\beta_1", output='character'), 
           parse=TRUE, size = 7)
```

.large[
If $\text{species} = \text{Adelie}, \hspace{1cm} \text{bill length} = \beta_0 + \varepsilon$

* $\beta_0$ = true mean bill length of Adelie penguins (in the population)

If $\text{species} = \text{Chinstrap}, \hspace{0.3cm} \text{bill length} = \beta_0 + \beta_1 + \varepsilon$
]

---

### Categorical predictor

```{r echo=F, message=F, warning=F, fig.align='center', fig.width=5, fig.height=3}

set.seed(1)
penguins_no_nas %>%
  ggplot(aes(x = species, y = bill_length_mm)) +
  geom_jitter(alpha = 0) +
  theme_classic() +
  geom_segment(aes(x = 0.6, y = 38, xend = 1.4, yend = 38), lwd=1.5, color="blue") + 
  geom_segment(aes(x = 1.6, y = 50, xend = 2.4, yend = 50), lwd=1.5, color="blue") + 
  geom_segment(aes(x = 2.6, y = 47, xend = 3.4, yend = 47), lwd=1.5, color="blue") + 
  labs(x = "Species", 
       y = "Bill length (mm)") +
  theme(text = element_text(size = 25)) +
  theme(axis.title.y = element_blank()) +
  scale_y_continuous(breaks = c(38, 50),
                     labels = c(TeX("$\\beta_0"), 
                                TeX("$\\beta_0 + \\beta_1"))) +
  geom_segment(aes(x = 1.45, y = 38,
                   xend = 1.45, yend = 50),
               arrow = arrow(length = unit(0.3, "cm"))) +
  annotate("text", x = 1.3, y = 44, 
           label = TeX("$\\beta_1", output='character'), 
           parse=TRUE, size = 7)
```

.large[
If $\text{species} = \text{Adelie}, \hspace{1cm} \text{bill length} = \beta_0 + \varepsilon$

If $\text{species} = \text{Chinstrap}, \hspace{0.3cm} \text{bill length} = \beta_0 + \beta_1 + \varepsilon$

.question[
What about Gentoo penguins?
]
]

---

### Categorical predictor

```{r echo=F, message=F, warning=F, fig.align='center', fig.width=5, fig.height=3}

set.seed(1)
penguins_no_nas %>%
  ggplot(aes(x = species, y = bill_length_mm)) +
  geom_jitter(alpha = 0) +
  theme_classic() +
  geom_segment(aes(x = 0.6, y = 38, xend = 1.4, yend = 38), lwd=1.5, color="blue") + 
  geom_segment(aes(x = 1.6, y = 50, xend = 2.4, yend = 50), lwd=1.5, color="blue") + 
  geom_segment(aes(x = 2.6, y = 47, xend = 3.4, yend = 47), lwd=1.5, color="blue") + 
  labs(x = "Species", 
       y = "Bill length (mm)") +
  theme(text = element_text(size = 25)) +
  theme(axis.title.y = element_blank()) +
  scale_y_continuous(breaks = c(38, 50, 47),
                     labels = c(TeX("$\\beta_0"), 
                                TeX("$\\beta_0 + \\beta_1"), 
                                TeX("$\\beta_0 + \\beta_2"))) +
  geom_segment(aes(x = 1.45, y = 38,
                   xend = 1.45, yend = 50),
               arrow = arrow(length = unit(0.3, "cm"))) +
  geom_segment(aes(x = 2.45, y = 38,
                   xend = 2.45, yend = 47),
               arrow = arrow(length = unit(0.3, "cm"))) +
  annotate("text", x = 1.3, y = 44, 
           label = TeX("$\\beta_1", output='character'), 
           parse=TRUE, size = 7) +
  annotate("text", x = 2.3, y = 43, 
           label = TeX("$\\beta_2", output='character'), 
           parse=TRUE, size = 7)
```

.large[
If $\text{species} = \text{Adelie}, \hspace{1cm} \text{bill length} = \beta_0 + \varepsilon$

If $\text{species} = \text{Chinstrap}, \hspace{0.3cm} \text{bill length} = \beta_0 + \beta_1 + \varepsilon$

If $\text{species} = \text{Gentoo}, \hspace{1cm} \text{bill length} = \beta_0 + \beta_2 + \varepsilon$
]

---

### Categorical predictor

```{r echo=F, message=F, warning=F, fig.align='center', fig.width=6, fig.height=4}

set.seed(1)
penguins_no_nas %>%
  ggplot(aes(x = species, y = bill_length_mm)) +
  geom_jitter(alpha = 0) +
  theme_classic() +
  geom_segment(aes(x = 0.6, y = 38, xend = 1.4, yend = 38), lwd=1.5, color="blue") + 
  geom_segment(aes(x = 1.6, y = 50, xend = 2.4, yend = 50), lwd=1.5, color="blue") + 
  geom_segment(aes(x = 2.6, y = 47, xend = 3.4, yend = 47), lwd=1.5, color="blue") + 
  labs(x = "Species", 
       y = "Bill length (mm)") +
  theme(text = element_text(size = 25)) +
  theme(axis.title.y = element_blank()) +
  scale_y_continuous(breaks = c(38, 50, 47),
                     labels = c(TeX("$\\beta_0"), 
                                TeX("$\\beta_0 + \\beta_1"), 
                                TeX("$\\beta_0 + \\beta_2"))) +
  geom_segment(aes(x = 1.45, y = 38,
                   xend = 1.45, yend = 50),
               arrow = arrow(length = unit(0.3, "cm"))) +
  geom_segment(aes(x = 2.45, y = 38,
                   xend = 2.45, yend = 47),
               arrow = arrow(length = unit(0.3, "cm"))) +
  annotate("text", x = 1.3, y = 44, 
           label = TeX("$\\beta_1", output='character'), 
           parse=TRUE, size = 7) +
  annotate("text", x = 2.3, y = 43, 
           label = TeX("$\\beta_2", output='character'), 
           parse=TRUE, size = 7)
```

.large[
$$\text{bill length} = \begin{cases} \beta_0 + \varepsilon & \text{species = Adelie} \\ \beta_0 + \beta_1 + \varepsilon & \text{species = Chinstrap} \\ \beta_0 + \beta_2 + \varepsilon & \text{species = Gentoo} \end{cases}$$
]

---

### Categorical predictor

```{r echo=F, message=F, warning=F, fig.align='center', fig.width=6, fig.height=4}

set.seed(1)
penguins_no_nas %>%
  ggplot(aes(x = species, y = bill_length_mm)) +
  geom_jitter(alpha = 0) +
  theme_classic() +
  geom_segment(aes(x = 0.6, y = 38, xend = 1.4, yend = 38), lwd=1.5, color="blue") + 
  geom_segment(aes(x = 1.6, y = 50, xend = 2.4, yend = 50), lwd=1.5, color="blue") + 
  geom_segment(aes(x = 2.6, y = 47, xend = 3.4, yend = 47), lwd=1.5, color="blue") + 
  labs(x = "Species", 
       y = "Bill length (mm)") +
  theme(text = element_text(size = 25)) +
  theme(axis.title.y = element_blank()) +
  scale_y_continuous(breaks = c(38, 50, 47),
                     labels = c(TeX("$\\beta_0"), 
                                TeX("$\\beta_0 + \\beta_1"), 
                                TeX("$\\beta_0 + \\beta_2"))) +
  geom_segment(aes(x = 1.45, y = 38,
                   xend = 1.45, yend = 50),
               arrow = arrow(length = unit(0.3, "cm"))) +
  geom_segment(aes(x = 2.45, y = 38,
                   xend = 2.45, yend = 47),
               arrow = arrow(length = unit(0.3, "cm"))) +
  annotate("text", x = 1.3, y = 44, 
           label = TeX("$\\beta_1", output='character'), 
           parse=TRUE, size = 7) +
  annotate("text", x = 2.3, y = 43, 
           label = TeX("$\\beta_2", output='character'), 
           parse=TRUE, size = 7)
```

.large[
$$\text{bill length} = \begin{cases} \beta_0 + \varepsilon & \text{species = Adelie} \\ \beta_0 + \beta_1 + \varepsilon & \text{species = Chinstrap} \\ \beta_0 + \beta_2 + \varepsilon & \text{species = Gentoo} \end{cases}$$
]

.large[
.question[
Can we write this more concisely?
]
]

---

### Indicator variables

.large[
$$\text{IsChinstrap} = \begin{cases} 0 & \text{species } \neq \text{Chinstrap} \\ 1 & \text{species = Chinstrap} \end{cases}$$

$$\text{IsGentoo} = \begin{cases} 0 & \text{species } \neq \text{Gentoo} \\ 1 & \text{species = Gentoo} \end{cases}$$
.question[
Do I need another indicator variable to represent Adelie penguins?
]
]

---

### Indicator variables

.large[
$$\text{IsChinstrap} = \begin{cases} 0 & \text{species } \neq \text{Chinstrap} \\ 1 & \text{species = Chinstrap} \end{cases}$$

$$\text{IsGentoo} = \begin{cases} 0 & \text{species } \neq \text{Gentoo} \\ 1 & \text{species = Gentoo} \end{cases}$$

Then:

| Species | IsChinstrap | IsGentoo |
| --- | --- | --- |
| Adelie | 0 | 0 |
| Chinstrap | 1 | 0 |
| Gentoo | 0 | 1 |
]

---

### Indicator variables

```{r echo=F, message=F, warning=F, fig.align='center', fig.width=6, fig.height=3.5}

set.seed(1)
penguins_no_nas %>%
  ggplot(aes(x = species, y = bill_length_mm)) +
  geom_jitter(alpha = 0) +
  theme_classic() +
  geom_segment(aes(x = 0.6, y = 38, xend = 1.4, yend = 38), lwd=1.5, color="blue") + 
  geom_segment(aes(x = 1.6, y = 50, xend = 2.4, yend = 50), lwd=1.5, color="blue") + 
  geom_segment(aes(x = 2.6, y = 47, xend = 3.4, yend = 47), lwd=1.5, color="blue") + 
  labs(x = "Species", 
       y = "Bill length (mm)") +
  theme(text = element_text(size = 25)) +
  theme(axis.title.y = element_blank()) +
  scale_y_continuous(breaks = c(38, 50, 47),
                     labels = c(TeX("$\\beta_0"), 
                                TeX("$\\beta_0 + \\beta_1"), 
                                TeX("$\\beta_0 + \\beta_2"))) +
  geom_segment(aes(x = 1.45, y = 38,
                   xend = 1.45, yend = 50),
               arrow = arrow(length = unit(0.3, "cm"))) +
  geom_segment(aes(x = 2.45, y = 38,
                   xend = 2.45, yend = 47),
               arrow = arrow(length = unit(0.3, "cm"))) +
  annotate("text", x = 1.3, y = 44, 
           label = TeX("$\\beta_1", output='character'), 
           parse=TRUE, size = 7) +
  annotate("text", x = 2.3, y = 43, 
           label = TeX("$\\beta_2", output='character'), 
           parse=TRUE, size = 7)
```

.large[
$$\text{bill length} = \begin{cases} \beta_0 + \varepsilon & \text{species = Adelie} \\ \beta_0 + \beta_1 + \varepsilon & \text{species = Chinstrap} \\ \beta_0 + \beta_2 + \varepsilon & \text{species = Gentoo} \end{cases}$$

Or, more concisely:

$$\text{bill length} = \beta_0 + \beta_1 \text{IsChinstrap} + \beta_2 \text{IsGentoo} + \varepsilon$$
]

---

### Estimated model

.large[
$$\widehat{\text{bill length}} = 38.82 + 10.01 \ \text{IsChinstrap} + 8.74 \ \text{IsGentoo}$$
]

.large[
.question[
* What is the estimated average bill length for each of the three species of penguin?
* There are 333 penguins in the dataset used to fit the model. How many degrees of freedom does the fitted model have?
]
]

---

### Estimated model

.large[
$$\widehat{\text{bill length}} =  38.82 + 10.01 \ \text{IsChinstrap} + 8.74 \ \text{IsGentoo}$$

| Species | $\widehat{\text{bill length}}$ |
| --- | --- |
| Adelie | 38.82 | 
| Chinstrap | 38.82 + 10.01 = 48.83 | 
| Gentoo | 38.82 + 8.74 = 47.56 |
]

---

### Estimated model

.large[
$$\widehat{\text{bill length}} =  38.82 + 10.01 \ \text{IsChinstrap} + 8.74 \ \text{IsGentoo}$$

| Species | $\widehat{\text{bill length}}$ |
| --- | --- |
| Adelie | 38.82 | 
| Chinstrap | 38.82 + 10.01 = 48.83 | 
| Gentoo | 38.82 + 8.74 = 47.56 |
]

.large[
**Degrees of freedom:** 333 - 3 = 330
* (Lose one degree of freedom for each parameter estimated)
]

---

### Fitting the model in R

.large[
R will automatically create indicator variables:

```{r}
lm(bill_length_mm ~ species, data = penguins_no_nas)
```

$$\widehat{\text{bill length}} =  38.82 + 10.01 \ \text{IsChinstrap} + 8.74 \ \text{IsGentoo}$$
]

---

### Class activity

.large[
* Practice regression with a single categorical predictor, with experimental data on fruit flies
* No need to submit this one on Canvas

[https://sta112-f25.github.io/class_activities/ca_15.html](https://sta112-f25.github.io/class_activities/ca_15.html)
]



