---
title: Exam 1 review day
output:
  xaringan::moon_reader:
    css: "lab-slides.css"
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
editor_options: 
  chunk_output_type: console
---

### Reminders and agenda

.large[
* Exam 1 on Friday
  * Bring a calculator!
  * Review questions and solutions on the course website
* Today:
  * Quick discussion about outliers activity from Monday
  * Exam 1 review
]

---

### Outliers

.large[
When can we *remove* potential outliers from the data?

* When the observation is a clear error (e.g., a negative value for height)
* When the observation comes from a different population than our population of interest (e.g., if we want to fit a model about dogs, then we can remove a cat from the data)

In all other situations:

* Report the results (estimated coefficients, confidence intervals, test statistics, etc.) both **with** AND **without** the potential outlier(s)
]

---

### Last time: exploring outliers

```{r, include=F}
library(tidyverse)
library(usdata)
data("urban_owner")

m1 <- lm(pct_owner_occupied ~ poppct_urban, data = urban_owner)
```

```{r, echo=F, message=F, warning=F, fig.align='center', fig.width=7, fig.height=5}
data.frame(studentized_resids = rstudent(m1),
         yhat = predict(m1)) |>
  ggplot(aes(x = yhat, y = studentized_resids)) +
  geom_point() +
  geom_abline(slope = 0, intercept = 0, 
              color = "blue", lwd=1.2) +
  geom_abline(slope = 0, intercept = c(-3, -2, 2, 3),
              color = "orange", lty = 2, lwd=1.2) +
  labs(x = "Predicted fraction owning their home",
       y = "Studentized residuals") +
  theme_bw() +
  theme(text = element_text(size = 20))
```

.large[
.question[
What point is the outlier here?
]
]

---

### Last time: exploring outliers

.large[
DC: Percent urban -- 100%, Percent owning home -- 42%

.question[
Is this observation a clear *error*?
]
]

---

### Last time: exploring outliers

.large[
DC: Percent urban -- 100%, Percent owning home -- 42%

.question[
Is this observation a clear *error*?
]

**No** -- these are believable numbers, given what we know about DC
]

---

### Last time: exploring outliers

.large[
DC: Percent urban -- 100%, Percent owning home -- 42%

.question[
Is this observation a clear *error*?
]

**No** -- these are believable numbers, given what we know about DC
]

---

### Last time: exploring outliers

.large[
DC: Percent urban -- 100%, Percent owning home -- 42%

.question[
Is this observation a clear *error*?
]

**No** -- these are believable numbers, given what we know about DC

.question[
Does DC belong to a different population than our population of interest?
]
]

---

### Last time: exploring outliers

.large[
DC: Percent urban -- 100%, Percent owning home -- 42%

.question[
Is this observation a clear *error*?
]

**No** -- these are believable numbers, given what we know about DC

.question[
Does DC belong to a different population than our population of interest?
]

**Maybe** -- I think you can make an argument either way
]

---

### Last time: exploring outliers

.large[
.question[
How should we handle DC?
]

* If we think DC does not belong in our population of interest, could remove
* Otherwise, report results with and without DC
]



