---
title: Nonlinear relationships and transformations
output:
  xaringan::moon_reader:
    css: "lab-slides.css"
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
editor_options: 
  chunk_output_type: console
---

### Last time: Class activity

.large[
$$\widehat{\text{LifeExpectancy}} = 4.95 + 7.20 \log(\text{GDP per capita})$$

.question[
What is the predicted price for GDP per capita of $20200?
]
]

---

### Class activity

.large[
$$\widehat{\text{LifeExpectancy}} = 4.95 + 7.20 \log(\text{GDP per capita})$$

* When GDP per capita = 20000: $\widehat{\text{LifeExpectancy}} = 76.255$
* When GDP per capita = 20200: $\widehat{\text{LifeExpectancy}} = 76.327$

Change:
]

---

### Class activity

.large[
$$\widehat{\log(\text{Price})} = 10.51 + 0.091 \ \text{Area}$$

* Price in $, Area in hundres of sq ft
]

.large[
.question[
What is the predicted price for a house of 2100 square feet?
]
]

---

### Class activity

.large[
$$\widehat{\log(\text{Price})} = 10.51 + 0.091 \ \text{Area}$$

* Price in $, Area in hundres of sq ft
* When Area = 20: $\widehat{\text{Price}} = 226386$
* When Area = 21: $\widehat{\text{Price}} = 247954$

Change:
]

---

### Class activity

.large[
$$\widehat{\log(\text{Price})} = 10.51 + 0.091 \ \text{Area}$$

* Price in $, Area in hundres of sq ft
* When Area = 20: $\widehat{\text{Price}} = 226386$
* When Area = 21: $\widehat{\text{Price}} = 247954$

Change: $\dfrac{247954}{226386} = 1.095 = e^{0.091}$
]


---

### More transformations


```{r, echo=F, message=F, warning=F, fig.width=10, fig.height=5}
library(tidyverse)
library(Stat2Data)
library(gridExtra)

data("SpeciesArea")

p1 <- SpeciesArea |>
  ggplot(aes(x = Area, y = Species)) +
  geom_point() +
  theme_bw() +
  labs(x = "Area (sq km)", y = "Number of species") +
  theme(text = element_text(size = 20)) +
  geom_smooth(se=F, method="lm")

m1 <- lm(Species ~ Area, data = SpeciesArea)

p2 <- data.frame(preds = predict(m1), 
                 resids = residuals(m1)) |>
  ggplot(aes(x = preds, y = resids)) +
  geom_point() +
  geom_abline(slope = 0, intercept = 0, color="blue") +
  theme_bw() +
  labs(x = "Predicted number of species", y = "Residuals", 
       title = "Residual plot") +
  theme(text = element_text(size = 20))

grid.arrange(p1, p2, ncol=2)
```

.large[
.question[
What transformations could we try here?
]
]

---

### Trying some transformations

```{r, echo=F, message=F, fig.width=8, fig.height=7, fig.align='center'}
p1 <- SpeciesArea |>
  ggplot(aes(x = log(Area), y = Species)) +
  geom_point() +
  theme_bw() +
  labs(x = "log(Area)", y = "Number of species") +
  theme(text = element_text(size = 20)) +
  geom_smooth(se=F, method="lm")

m1 <- lm(Species ~ log(Area), data = SpeciesArea)

p2 <- data.frame(preds = predict(m1), 
                 resids = residuals(m1)) |>
  ggplot(aes(x = preds, y = resids)) +
  geom_point() +
  geom_abline(slope = 0, intercept = 0, color="blue") +
  theme_bw() +
  labs(x = "Predicted number of species", y = "Residuals") +
  theme(text = element_text(size = 20))


p3 <- SpeciesArea |>
  ggplot(aes(x = sqrt(Area), y = Species)) +
  geom_point() +
  theme_bw() +
  labs(x = "sqrt(Area)", y = "Number of species") +
  theme(text = element_text(size = 20)) +
  geom_smooth(se=F, method="lm")

m1 <- lm(Species ~ sqrt(Area), data = SpeciesArea)

p4 <- data.frame(preds = predict(m1), 
                 resids = residuals(m1)) |>
  ggplot(aes(x = preds, y = resids)) +
  geom_point() +
  geom_abline(slope = 0, intercept = 0, color="blue") +
  theme_bw() +
  labs(x = "Predicted number of species", y = "Residuals") +
  theme(text = element_text(size = 20))



grid.arrange(p1, p2, p3, p4, ncol=2)
```

---

### Transform both predictor and response?

.large[
$$\log(Species) = \beta_0 + \beta_1 \log(Area) + \varepsilon$$
]

```{r, echo=F, message=F, warning=F, fig.width=10, fig.height=5}
p1 <- SpeciesArea |>
  ggplot(aes(x = log(Area), y = log(Species))) +
  geom_point() +
  theme_bw() +
  labs(x = "log(Area)", y = "log(Number of species)") +
  theme(text = element_text(size = 20)) +
  geom_smooth(se=F, method="lm")

m1 <- lm(log(Species) ~ log(Area), data = SpeciesArea)

p2 <- data.frame(preds = predict(m1), 
                 resids = residuals(m1)) |>
  ggplot(aes(x = preds, y = resids)) +
  geom_point() +
  geom_abline(slope = 0, intercept = 0, color="blue") +
  theme_bw() +
  labs(x = "Predicted number of species", y = "Residuals") +
  theme(text = element_text(size = 20))

grid.arrange(p1, p2, ncol=2)
```

---

### Interpreting the fitted model

.large[
$$\widehat{\log(Species)} = 1.625 + 0.246 \log (Area)$$
]

---

### Summary of model interpretation so far

$\widehat{y} = \widehat{\beta}_0 + \widehat{\beta}_1 x$

* A one-unit increase in $x$ is associated with a change in $y$ of $\widehat{\beta}_1$, on average

$\widehat{y} = \widehat{\beta}_0 + \widehat{\beta}_1 log(x)$

* A 1% increase in $x$ is associated with a change in $y$ of $\widehat{\beta}_1/100$, on average

$\widehat{log(y)} = \widehat{\beta}_0 + \widehat{\beta}_1 x$

* A one-unit increase in $x$ is associated with a change in $y$ by a **factor** of $e^{\widehat{\beta}_1}$, on average

$\widehat{log(y)} = \widehat{\beta}_0 + \widehat{\beta}_1 log(x)$

* A 1% increase in $x$ is associated with a change in $y$ by a **factor** of $e^{\widehat{\beta}_1}$, on average


---

### Kentucky Derby data

.large[
Data on the Kentucky Derby winner from each year between 1875 and 2012. Variables include

* `year`: year of race
* `speed`: speed of winning horse (mph)
* `time`: winning time (seconds)
* `condition`: condition of the track on the day of the race
]

---

### EDA

```{r echo=F, message=F, warning=F, fig.align='center', fig.width=7, fig.height=5}
library(tidyverse)
derby <- read.delim("https://sta112-s22.github.io/slides/KYDerby17.txt")

derby %>%
  ggplot(aes(x = Year, y = Time)) +
  geom_point(size = 2.5) +
  geom_line() +
  theme_bw() +
  theme(text = element_text(size = 25))
```

.large[
.question[
What do you notice about the winning times?
]
]

---

### EDA

```{r echo=F, message=F, warning=F, fig.align='center', fig.width=7, fig.height=5}
library(tidyverse)
derby <- read.delim("https://sta112-s22.github.io/slides/KYDerby17.txt")

derby %>%
  ggplot(aes(x = Year, y = Time)) +
  geom_point(size = 2.5) +
  geom_line() +
  theme_bw() +
  theme(text = element_text(size = 25))

derby <- derby |> filter(Year > 1900)
```

.large[
.question[
What do you notice about the winning times?
]
]

.large[
There's a big drop around 1900, then they decrease gradually.
]

---

### EDA

```{r echo=F, message=F, warning=F, fig.align='center', fig.width=7, fig.height=5}

derby %>%
  ggplot(aes(x = Year, y = speed)) +
  geom_point(size = 2.5) +
  theme_bw() +
  theme(text = element_text(size = 25))
```

.large[
.question[
How would you describe the relationship?
]
]

---

### Fitting a linear regression

.large[
$$\text{Speed} = \beta_0 + \beta_1 \ \text{Year} + \varepsilon$$
]

```{r echo=F, message=F, warning=F, fig.align='center', fig.width=10, fig.height=5}

p1 <- derby %>%
  ggplot(aes(x = Year, y = speed)) +
  geom_point(size = 2.5) +
  geom_smooth(se=F, method="lm") +
  theme_bw() +
  theme(text = element_text(size = 20))

derby_lm <- lm(speed ~ Year, data = derby)

p2 <- derby %>%
  mutate(pred = predict(derby_lm),
         resid = residuals(derby_lm)) %>%
  ggplot(aes(x = pred, y = resid)) +
  geom_point(size = 2.5) +
  geom_abline(slope = 0, intercept = 0, color = "blue", lwd=1.2) +
  labs(x = "Predicted speed", y = "Residual") +
  theme_bw() +
  theme(text = element_text(size = 20))

grid.arrange(p1, p2, ncol=2)
```

---

### Polynomial transformations

.large[
Another possible transformation is a **polynomial**.

.question[
**Definition:** A polynomial transformation of order $k$ is a linear combination of the terms $X$, $X^2$, ..., $X^k$:
$$Y = \beta_0 + \beta_1 X + \beta_2 X^2 + \cdots + \beta_k X^k + \varepsilon$$
]
]

.large[
* Example: If we use a polynomial of order 3, our model is 
$$Y = \beta_0 + \beta_1 X + \beta_2 X^2 + \beta_3 X^3 + \varepsilon$$
]

---

### Polynomial transformations

```{r echo=F, message=F, warning=F, fig.align='center', fig.width=7, fig.height=5}

derby %>%
  ggplot(aes(x = Year, y = speed)) +
  geom_point(size = 2.5) +
  theme_bw() +
  theme(text = element_text(size = 25))
```

.large[
.question[
What order polynomial do you think we should use?
]
]

---

### Polynomial transformation

.large[
Let's try a polynomial of order 2:

$$\text{Speed} = \beta_0 + \beta_1 \ \text{Year} +  \beta_2 \ \text{Year}^2 + \varepsilon$$
]

```{r echo=F, message=F, warning=F, fig.align='center', fig.width=7, fig.height=5}

derby %>%
  ggplot(aes(x = Year, y = speed)) +
  geom_point(size = 2.5) +
  geom_smooth(se=F, method="lm", formula = y ~ I(x) + I(x^2), lwd=1.2) +
  theme_bw() +
  theme(text = element_text(size = 25))
```

---

### Diagnostic plots

.large[
We use residual plots to check the model assumptions after transforming our predictor, just like with other transformations.
]

```{r echo=F, message=F, warning=F, fig.align='center', fig.width=6, fig.height=4}

derby_lm <- lm(speed ~ I(Year) + I(Year^2), data = derby)

derby %>%
  mutate(pred = predict(derby_lm),
         resid = residuals(derby_lm)) %>%
  ggplot(aes(x = pred, y = resid)) +
  geom_point(size = 2.5) +
  geom_abline(slope = 0, intercept = 0, color = "blue", lwd=1.2) +
  labs(x = "Predicted speed", y = "Residual") +
  theme_bw() +
  theme(text = element_text(size = 25))
```

.large[
.question[
How does the shape assumption look now?
]
]

---

### Fitting the model in R

.large[
```{r eval=F}
derby_lm <- lm(speed ~ poly(Year, 2, raw=TRUE), 
               data = derby)
```
]

.large[
* `poly` means "fit a polynomial"
* `poly(..., 2, ...)` specifies a second-order polynomial (quadratic)
]

---

### Fitting the model in R

.large[
```{r}
derby_lm <- lm(speed ~ poly(Year, 2, raw=TRUE), 
               data = derby)
derby_lm
```
]

.large[
$$\widehat{\text{Speed}} = -1120 + 1.16 \ \text{Year} - 0.00029 \ \text{Year}^2$$
]

---

### Interpreting the fitted model

.large[
$$\widehat{\text{Speed}} = -1120 + 1.16 \ \text{Year} - 0.00029 \ \text{Year}^2$$
]

.large[
.question[
How would I interpret the intercept of the fitted polynomial regression model?
]
]

---

### Interpreting the fitted model

.large[
$$\widehat{\text{Speed}} = -1120 + 1.16 \ \text{Year} - 0.00029 \ \text{Year}^2$$
]

.large[
.question[
How would I interpret the intercept of the fitted polynomial regression model?
]
]

.large[
If Year = 0, we predict a speed of -1120 mph (doesn't make much sense, since year is nowhere near 0)
]

---

### Interpreting the fitted model

.large[
$$\widehat{\text{Speed}} = -1120 + 1.16 \ \text{Year} - 0.00029 \ \text{Year}^2$$
]

.large[
.question[
How would I interpret the terms on $\text{Year}$ and $\text{Year}^2$?
]
]

---

### Interpreting the fitted model

.large[
$$\widehat{\text{Speed}} = -1120 + 1.16 \ \text{Year} - 0.00029 \ \text{Year}^2$$
]

.large[
.question[
How would I interpret the terms on $\text{Year}$ and $\text{Year}^2$?
]
]

.large[
They're kind of difficult to interpret -- if I change $\text{Year}$, I also change $\text{Year}^2$
]

---

### Interpreting the fitted model

.large[
$$\widehat{\text{Speed}} = -1120 + 1.16 \ \text{Year} - 0.00029 \ \text{Year}^2$$
]

.large[
.question[
Can I interpret the *sign* on $\text{Year}^2$?
]
]

---

### Interpreting the fitted model

.large[
$$\widehat{\text{Speed}} = -1120 + 1.16 \ \text{Year} - 0.00029 \ \text{Year}^2$$
]

.large[
.question[
Can I interpret the *sign* on $\text{Year}^2$?
]
]

.large[
Yes (for a polynomial of order 2 specifically). The sign on $\text{Year}^2$ tells us about the concavity of the relationship.
]

---

### Class activity

.large[
[https://sta112-f25.github.io/class_activities/ca_11.html](https://sta112-f25.github.io/class_activities/ca_11.html)
]

.large[
* Experiment with polynomial regression
* Submit HTML at end of class
]






