---
title: Assumptions, diagnostics, and multicollinearity
output:
  xaringan::moon_reader:
    css: "lab-slides.css"
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
editor_options: 
  chunk_output_type: console
---

```{r include=F}
library(tidyverse)
library(Stat2Data)
data(Pines)
```


### Last time: Variance inflation factors

.large[
**Motivation:** High multicollinearity means a predictor can be estimated well by some combination of other predictors

**Variance inflation factor:** Let $X_i$ be a predictor in the model, and $R_i^2$ the coefficient of determination for a model estimating $X_i$ from the other predictors. The *variance inflation factor* of $X_i$ is

$$VIF_i = \dfrac{1}{1 - R_i^2}$$
* Higher VIF means more multicollinearity
* Rule of thumb: worried when $VIF_i > 5$ (i.e. $R_i^2 > 0.8$ )
]

---

### Last time: Class activity

.large[
* $\widehat{\text{Hgt97}} = 40.59 + 1.10 \ \text{Hgt96} \hspace{1cm} R^2_{adj} =$
]

---

### Class activity

.large[
* $\widehat{\text{Hgt97}} = 40.59 + 1.10 \ \text{Hgt96} \hspace{1cm} R^2_{adj} = 0.949$
* $\widehat{\text{Hgt97}} = 40.24 − 1.26 \ \text{Diam96} + 1.12 \ \text{Hgt96} \hspace{1cm} R^2_{adj} =$
]

---

### Class activity

.large[
* $\widehat{\text{Hgt97}} = 40.59 + 1.10 \ \text{Hgt96} \hspace{1cm} R^2_{adj} = 0.949$
* $\widehat{\text{Hgt97}} = 40.24 − 1.26 \ \text{Diam96} + 1.12 \ \text{Hgt96} \hspace{1cm} R^2_{adj} = 0.942$

.question[
Why doesn't $R^2_{adj}$ increase when I add diameter to the model?
]
]

---

### Class activity

.large[
.question[
Why doesn't $R^2_{adj}$ increase when I add diameter to the model?
]
]

```{r echo=F, message=F, warning=F, fig.align='center', fig.width=7, fig.height=5}
data("Pines")
Pines %>%
  ggplot(aes(x = Diam96, y = Hgt96)) +
  geom_point(size=2) +
  theme_bw() +
  theme(text = element_text(size = 25))
```

.large[
Because diameter and height in 1996 are so correlated.
]

---

### Class activity

.large[
$$\widehat{\text{Hgt97}} = 40.24 − 1.26 \ \text{Diam96} + 1.12 \ \text{Hgt96}$$
.question[
What is the variance inflation factor for Diam96?
]
]

---

### Class activity

.large[
$$\widehat{\text{Hgt97}} = 40.24 − 1.26 \ \text{Diam96} + 1.12 \ \text{Hgt96}$$
.question[
What is the variance inflation factor for Diam96?
]
]

.large[
```{r}
m1 <- lm(Diam96 ~ Hgt96, data = Pines)
summary(m1)$r.squared
```

$VIF = \dfrac{1}{1 - 0.813} = 5.35$
]

---

### Why is multicollinearity a problem?

.large[
* Inflates variability of estimated coefficients
* Interpretation of individual terms is difficult

$$\widehat{\sqrt{\text{MDs}}} = 3.58 + 2.58 \ \text{Hospitals} + 0.012 \ \text{Beds}$$

**Usual interpretation:** Holding Beds fixed, an increase of 1 hospital is associated with an increase of 2.58 units in $\sqrt{\text{MDs}}$

**Problem:** When Beds and Hospitals are highly correlated, doesn't make sense to fix one and change the other. (Can't really change the number of hospitals without changing the number of beds)

]


---

### Example data

.large[

* `MED_EARNINGS`: median earnings for graduates from each university
* `CONTROL`: whether the school is public or private
* `SATVRMID`: midpoint of SAT critical reading scores of students attending the school
* `SATMTMID`: midpoint of SAT math scores
* `SATWRMID`: midpoint of SAT writing scores
* `ACTCMMID`: midpoint of the ACT cumulative scores
* `ACTENMID`: midpoint of ACT English scores
* `ACTMTMID`: midpoint of ACT math scores
* `+` others...

.question[
Which variables might exhibit multicollinearity?
]
]

---

### Example data

.large[

* `MED_EARNINGS`: median earnings for graduates from each university
* `CONTROL`: whether the school is public or private
* `SATVRMID`: midpoint of SAT critical reading scores of students attending the school
* `SATMTMID`: midpoint of SAT math scores
* `SATWRMID`: midpoint of SAT writing scores
* `ACTCMMID`: midpoint of the ACT cumulative scores
* `ACTENMID`: midpoint of ACT English scores
* `ACTMTMID`: midpoint of ACT math scores
* `+` others...

There is probably multicollinearity with all the test score variables
]

---

### Example data

.large[

* `MED_EARNINGS`: median earnings for graduates from each university
* `CONTROL`: whether the school is public or private
* `SATVRMID`: midpoint of SAT critical reading scores of students attending the school
* `SATMTMID`: midpoint of SAT math scores
* `SATWRMID`: midpoint of SAT writing scores
* `ACTCMMID`: midpoint of the ACT cumulative scores
* `ACTENMID`: midpoint of ACT English scores
* `ACTMTMID`: midpoint of ACT math scores
* `+` others...

.question[
How can we visually investigate correlated predictors?
]
]

---

### Example data

.pull-left[
```{r, echo=F, message=F, fig.width=6, fig.height=4}

scorecard <- read_csv("~/Documents/Teaching/sta112/sta112-s22.github.io/projects/scorecard.csv")

scorecard <- scorecard |>
  rename(MED_EARNINGS = MD_EARN_WNE_P10)

scorecard |>
  ggplot(aes(x = SATVRMID, y = SATMTMID)) +
  geom_point() +
  theme_bw()
```

.large[
$r = 0.95$
]
]

.pull-right[

```{r, echo=F, message=F, fig.width=6, fig.height=4}

scorecard |>
  ggplot(aes(x = ACTMTMID, y = SATMTMID)) +
  geom_point() +
  theme_bw()
```

.large[
$r = 0.94$
]
]

---

### VIFs in R

.large[
Fortunately, R makes it easy to calculate VIFs:

```{r message=F, warning=F}
library(car)
m1 <- lm(MED_EARNINGS ~ SATVRMID + SATMTMID + 
           SATWRMID + ACTCMMID + ACTENMID + 
           ACTMTMID + ACTWRMID,
         data = scorecard)
vif(m1)
```

* The `vif` function (from the `car` package) calculates a VIF for each predictor
* The variance inflation factor for SAT verbal score is 29.78, e.g.
]

---

### Handling multicollinearity

.large[
We have a few options for handling multicollinearity:

* Remove variables
  * Example: just pick one of the test score variables, and discard the rest, since they are so highly correlated
]

---

### Handling multicollinearity

.large[
We have a few options for handling multicollinearity:

* Remove variables
  * Example: just pick one of the test score variables, and discard the rest, since they are so highly correlated
* Create a new variable
  * Example: take an average or sum of the test score variables, instead of using all of them individually
]

---

### Handling multicollinearity

.large[
We have a few options for handling multicollinearity:

* Remove variables
  * Example: just pick one of the test score variables, and discard the rest, since they are so highly correlated
* Create a new variable
  * Example: take an average or sum of the test score variables, instead of using all of them individually
* Ignore it
  * If we don't care about interpreting coefficients or doing inference, multicollinearity doesn't matter as much
]








