---
title: Least-Squares Regression
output:
  xaringan::moon_reader:
    css: "lab-slides.css"
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
editor_options: 
  chunk_output_type: console
---

### Agenda

.large[
* Today: 
    * How do we estimate coefficients in a linear regression model?
    * More on prediction
* Later this week
    * Prediction and correlation
    * Activity: practicing statistical communication
]


---


### Recap: regression

.large[
**Regression:** model the relationship between predictor $x$ and response $y$

**Regression model:** $y = f(x) + \varepsilon$

* $f$ = systematic information that $x$ provides about $y$
* $\varepsilon$ = random noise

**Linear regression:** $y = \beta_0 + \beta_1 x + \varepsilon$

]

---

### Recap: linear regression

.large[
**Observed data (sample):** Weight and wing length for 116 sparrows
]

```{r echo=F, message=F, warning=F, fig.align='center', fig.width=6, fig.height=4}
library(Stat2Data)
library(tidyverse)
data(Sparrows)

Sparrows %>%
  ggplot(aes(x = Weight, y = WingLength)) +
  geom_point(size=2.5) +
  labs(x = "Weight (g)", 
       y = "Wing Length (mm)") +
  theme_bw() +
  theme(text = element_text(size=25))
```

.large[
**Regression model (population):**

.center[
$\text{WingLength} = \beta_0 + \beta_1 \text{Weight} + \varepsilon$
]
]

---

### Recap: linear regression

```{r echo=F, message=F, warning=F, fig.align='center', fig.width=6, fig.height=4}
library(Stat2Data)
library(tidyverse)
data(Sparrows)

Sparrows %>%
  ggplot(aes(x = Weight, y = WingLength)) +
  geom_point(size=2.5) +
  labs(x = "Weight (g)", 
       y = "Wing Length (mm)") +
  theme_bw() +
  theme(text = element_text(size=25))
```

.large[

**Estimated regression line (from sample):**

.center[
$\widehat{\text{WingLength}} = 8.76 + 1.31 \ \text{Weight}$
]
]

.large[
* $\widehat{\beta}_0 = 8.76$
* $\widehat{\beta}_1 = 1.31$
]
    
---

### Recap: interpreting regression coefficients

.large[
.center[
$\widehat{\text{WingLength}} = 8.76 + 1.31 \ \text{Weight}$
]
]

.large[
**Interpreting $\widehat{\beta}_0$:** If weight = 0g, we predict a wing length of 8.76mm.

**Interpreting $\widehat{\beta}_1$:** An increase of 1g in weight is associated with an increase of 1.31mm, on average, in wing length
]

---

### Concept check

.large[
Interested in the relationship between mileage and price for used Honda Accords. We will use linear regression.

.pull-left[

```{r, echo=F, fig.align='center', fig.width=6, fig.height=4}
data("AccordPrice")
AccordPrice %>%
  ggplot(aes(x = Mileage, y = Price)) +
  geom_point(size=2.5) +
  labs(x = "Miles (in 1000's)",
       y = "Price (in $1000's)") +
  theme_bw() +
  theme(text = element_text(size=25))
```
]


.pull-right[
Write the population regression model:

.abox[
(A) $\text{price} = \beta_0 + \beta_1 \text{miles}$
]

.bbox[
(B) $\text{price} = \beta_0 + \beta_1 \text{miles} + \varepsilon$
]

.cbox[
(C) $\widehat{\text{price}} = \widehat{\beta}_0 + \widehat{\beta}_1 \text{miles}$
]

.dbox[
(D) $\widehat{\text{price}} = \widehat{\beta}_0 + \widehat{\beta}_1 \text{miles} + \varepsilon$
]

]
]

---

### Concept check

.large[
Interested in the relationship between mileage and price for used Honda Accords. We will use linear regression.

.pull-left[

```{r, echo=F, fig.align='center', fig.width=6, fig.height=4}
data("AccordPrice")
AccordPrice %>%
  ggplot(aes(x = Mileage, y = Price)) +
  geom_point(size=2.5) +
  labs(x = "Miles (in 1000's)",
       y = "Price (in $1000's)") +
  theme_bw() +
  theme(text = element_text(size=25))
```
]


.pull-right[
Write the population regression model:

.abox[
(A) $\text{price} = \beta_0 + \beta_1 \text{miles}$
]

.bbox[
(B) $\text{price} = \beta_0 + \beta_1 \text{miles} + \varepsilon$
]

.cbox[
(C) $\widehat{\text{price}} = \widehat{\beta}_0 + \widehat{\beta}_1 \text{miles}$
]

.dbox[
(D) $\widehat{\text{price}} = \widehat{\beta}_0 + \widehat{\beta}_1 \text{miles} + \varepsilon$
]

]
]

.large[
**Answer:** (B)
]

---

### Linear regression

```{r echo=F, message=F, fig.align='center', fig.width = 6, fig.height=4}
AccordPrice %>%
  ggplot(aes(x = Mileage, y = Price)) +
  geom_point(size=2.5) +
  geom_smooth(se=F, method="lm") +
  labs(x = "Miles (in 1000's)",
       y = "Price (in $1000's)") +
  theme_bw() +
  theme(text = element_text(size=25))
```

.large[
**Estimated line** (from sample): $\widehat{\text{price}} = 20.81 - 0.12 \ \text{miles}$

How do we calculate $\widehat{\beta}_0$ and $\widehat{\beta}_1$?
]

---

### Choosing between different lines

```{r echo=F, message=F, fig.align='center', fig.width = 12, fig.height=4}
library(patchwork)
p1 <- AccordPrice %>%
  ggplot(aes(x = Mileage, y = Price)) +
  geom_point(size=2.5) +
  geom_abline(slope = -0.1, intercept = 27, color = "blue",
              lwd = 1.5) +
  labs(x = "Miles (in 1000's)",
       y = "Price (in $1000's)") +
  theme_bw() +
  theme(text = element_text(size=25))

p2 <- AccordPrice %>%
  ggplot(aes(x = Mileage, y = Price)) +
  geom_point(size=2.5) +
  geom_abline(slope = -0.1, intercept = 20, color = "blue",
              lwd = 1.5) +
  labs(x = "Miles (in 1000's)",
       y = "Price (in $1000's)") +
  theme_bw() +
  theme(text = element_text(size=25))

p3 <- AccordPrice %>%
  ggplot(aes(x = Mileage, y = Price)) +
  geom_point(size=2.5) +
  geom_abline(slope = -0.02, intercept = 15, color = "blue",
              lwd = 1.5) +
  labs(x = "Miles (in 1000's)",
       y = "Price (in $1000's)") +
  theme_bw() +
  theme(text = element_text(size=25))

p1 + p2 + p3
```

.large[
.question[
Which one looks like the best choice?
]
]

---

### Choosing between different lines

```{r echo=F, message=F, fig.align='center', fig.width = 12, fig.height=4}
library(patchwork)
p1 <- AccordPrice %>%
  ggplot(aes(x = Mileage, y = Price)) +
  geom_point(size=2.5) +
  geom_abline(slope = -0.1, intercept = 27, color = "blue",
              lwd = 1.5) +
  labs(x = "Miles (in 1000's)",
       y = "Price (in $1000's)") +
  theme_bw() +
  theme(text = element_text(size=25))

p2 <- AccordPrice %>%
  ggplot(aes(x = Mileage, y = Price)) +
  geom_point(size=2.5) +
  geom_abline(slope = -0.1, intercept = 20, color = "blue",
              lwd = 1.5) +
  labs(x = "Miles (in 1000's)",
       y = "Price (in $1000's)") +
  theme_bw() +
  theme(text = element_text(size=25))

p3 <- AccordPrice %>%
  ggplot(aes(x = Mileage, y = Price)) +
  geom_point(size=2.5) +
  geom_abline(slope = -0.02, intercept = 15, color = "blue",
              lwd = 1.5) +
  labs(x = "Miles (in 1000's)",
       y = "Price (in $1000's)") +
  theme_bw() +
  theme(text = element_text(size=25))

p1 + p2 + p3
```

.large[
.question[
Which one looks like the best choice?
]
]

.large[
Intuitively, want the line to "go through" the points. Can we formalize this?
]

---

### Residuals

```{r echo=F, message=F, warning=F, fig.align='center', fig.width = 6, fig.height=4}
AccordPrice %>%
  ggplot(aes(x = Mileage, y = Price)) +
  geom_segment(aes(x = 20.8, y = 17.92, 
                   xend = 20.8, yend = 22.8),
                   color = "red", lwd=1.5) +
  geom_point(size=2.5) +
  geom_abline(slope = -0.1, intercept = 20, color = "blue",
              lwd = 1.5) +
  labs(x = "Miles", 
       y = "Price") +
  theme_bw() +
  theme(text = element_text(size=25))
```

.large[
Car with mileage = 20.8

* Prediction: $\widehat{Price} = 18$
* Observed: $Price = 22.8$

Residual: $Price - \widehat{Price} = 22.8 - 18 = 4.8$
]

---

### Residuals

```{r echo=F, message=F, warning=F, fig.align='center', fig.width = 6, fig.height=3.5}
AccordPrice %>%
  ggplot(aes(x = Mileage, y = Price)) +
  geom_segment(aes(x = 20.8, y = 17.92, 
                   xend = 20.8, yend = 22.8),
                   color = "red", lwd=1.2) +
  geom_segment(aes(x = 89.4, y = 3.5, 
                   xend = 89.4, yend = 11.06),
                   color = "red", lwd=1.2) +
  geom_point(size=2) +
  geom_abline(slope = -0.1, intercept = 20, color = "blue",
              lwd = 1.2) +
  labs(x = "Miles", 
       y = "Price") +
  theme_bw() +
  theme(text = element_text(size=20))
```

.pull-left[
.large[
* Prediction: $\widehat{Price} = 18$
* Observed: $Price = 22.8$

$Price - \widehat{Price} = 4.8$
]
]

.pull-right[
.large[
* Prediction: $\widehat{Price} = 11.06$
* Observed: $Price = 3.5$

$Price - \widehat{Price} = -7.56$
]
]

.large[
.question[
Which of these predictions is closer?
]
]

---

### Residuals

```{r echo=F, message=F, warning=F, fig.align='center', fig.width = 6, fig.height=3.5}
AccordPrice %>%
  ggplot(aes(x = Mileage, y = Price)) +
  geom_segment(aes(x = 20.8, y = 17.92, 
                   xend = 20.8, yend = 22.8),
                   color = "red", lwd=1.2) +
  geom_segment(aes(x = 89.4, y = 3.5, 
                   xend = 89.4, yend = 11.06),
                   color = "red", lwd=1.2) +
  geom_point(size=2) +
  geom_abline(slope = -0.1, intercept = 20, color = "blue",
              lwd = 1.2) +
  labs(x = "Miles", 
       y = "Price") +
  theme_bw() +
  theme(text = element_text(size=20))
```

.pull-left[
.large[
* Prediction: $\widehat{Price} = 18$
* Observed: $Price = 22.8$

$Price - \widehat{Price} = 4.8$
]
]

.pull-right[
.large[
* Prediction: $\widehat{Price} = 11.06$
* Observed: $Price = 3.5$

$Price - \widehat{Price} = -7.56$
]
]

.large[
.question[
What does a *positive* residual mean?
]
]

---

### Residuals

```{r echo=F, message=F, warning=F, fig.align='center', fig.width = 6, fig.height=3.5}
AccordPrice %>%
  ggplot(aes(x = Mileage, y = Price)) +
  geom_segment(aes(x = 20.8, y = 17.92, 
                   xend = 20.8, yend = 22.8),
                   color = "red", lwd=1.2) +
  geom_segment(aes(x = 89.4, y = 3.5, 
                   xend = 89.4, yend = 11.06),
                   color = "red", lwd=1.2) +
  geom_point(size=2) +
  geom_abline(slope = -0.1, intercept = 20, color = "blue",
              lwd = 1.2) +
  labs(x = "Miles", 
       y = "Price") +
  theme_bw() +
  theme(text = element_text(size=20))
```

.pull-left[
.large[
* Prediction: $\widehat{Price} = 18$
* Observed: $Price = 22.8$

$Price - \widehat{Price} = 4.8$
]
]

.pull-right[
.large[
* Prediction: $\widehat{Price} = 11.06$
* Observed: $Price = 3.5$

$Price - \widehat{Price} = -7.56$
]
]

.large[
.question[
What does a *negative* residual mean?
]
]

---

### Residuals

```{r echo=F, message=F, fig.align='center', fig.width = 12, fig.height=4}
p1 <- AccordPrice %>%
  ggplot(aes(x = Mileage, y = Price)) +
  geom_segment(aes(xend = Mileage,
                   yend = 27 - 0.1*Mileage),
               color = "red",
               lwd=1.5) +
  geom_point(size=2.5) +
  geom_abline(slope = -0.1, intercept = 27, color = "blue",
              lwd = 1.5) +
  labs(x = "Miles", 
       y = "Price") +
  theme_bw() +
  theme(text = element_text(size=25))

p2 <- AccordPrice %>%
  ggplot(aes(x = Mileage, y = Price)) +
  geom_segment(aes(xend = Mileage,
                   yend = 20 - 0.1*Mileage),
               color = "red",
               lwd=1.5) +
  geom_point(size=2.5) +
  geom_abline(slope = -0.1, intercept = 20, color = "blue",
              lwd = 1.5) +
  labs(x = "Miles", 
       y = "Price") +
  theme_bw() +
  theme(text = element_text(size=25))

p3 <- AccordPrice %>%
  ggplot(aes(x = Mileage, y = Price)) +
  geom_segment(aes(xend = Mileage,
                   yend = 15 - 0.02*Mileage),
               color = "red",
               lwd=1.5) +
  geom_point(size=2.5) +
  geom_abline(slope = -0.02, intercept = 15, color = "blue",
              lwd = 1.5) +
  labs(x = "Miles", 
       y = "Price") +
  theme_bw() +
  theme(text = element_text(size=25))

p1 + p2 + p3
```

.large[
.question[
Which of these lines looks like the best fit? Why?
]
]

---

### Sum of the squared residuals
               
```{r echo=F, message=F, fig.align='center', fig.width = 12, fig.height=3}
p1 <- AccordPrice %>%
  mutate(fitted = 27 - 0.1*Mileage) %>%
  ggplot(aes(x = Mileage, y = Price)) +
  geom_rect(aes(xmin = Mileage, 
                xmax = Mileage + Price - fitted,
                ymin = Price, ymax = fitted), 
            fill = "red", color = "red", alpha = 0.2) +
  geom_point(size=2.5) +
  geom_abline(slope = -0.1, intercept = 27, color = "blue",
              lwd = 1.5) +
  labs(x = "Miles", 
       y = "Price",
       title = "SSE = 1871.17") +
  theme_bw() +
  theme(text = element_text(size=25))

p2 <- AccordPrice %>%
  mutate(fitted = 20 - 0.1*Mileage) %>%
  ggplot(aes(x = Mileage, y = Price)) +
  geom_rect(aes(xmin = Mileage, 
                xmax = Mileage + Price - fitted,
                ymin = Price, ymax = fitted), 
            fill = "red", color = "red", alpha = 0.2) +
  geom_point(size=2.5) +
  geom_abline(slope = -0.1, intercept = 20, color = "blue",
              lwd = 1.5) +
  labs(x = "Miles", 
       y = "Price",
       title = "SSE = 287.49") +
  theme_bw() +
  theme(text = element_text(size=25))

p3 <- AccordPrice %>%
  mutate(fitted = 15 - 0.02*Mileage) %>%
  ggplot(aes(x = Mileage, y = Price)) +
  geom_rect(aes(xmin = Mileage, 
                xmax = Mileage + Price - fitted,
                ymin = Price, ymax = fitted), 
            fill = "red", color = "red", alpha = 0.2) +
  geom_point(size=2.5) +
  geom_abline(slope = -0.02, intercept = 15, color = "blue",
              lwd = 1.5) +
  labs(x = "Miles", 
       y = "Price",
       title = "SSE = 747.78") +
  theme_bw() +
  theme(text = element_text(size=25))

p1 + p2 + p3
```

.large[
Sum of squared residuals (SSE) = $\sum_{i=1}^n (y_i - \widehat{y}_i)^2$
]

---

### Least Squares

```{r echo=F, message=F, fig.align='center', fig.width = 12, fig.height=3}
p1 <- AccordPrice %>%
  mutate(fitted = 27 - 0.1*Mileage) %>%
  ggplot(aes(x = Mileage, y = Price)) +
  geom_rect(aes(xmin = Mileage, 
                xmax = Mileage + Price - fitted,
                ymin = Price, ymax = fitted), 
            fill = "red", color = "red", alpha = 0.2) +
  geom_point(size=2.5) +
  geom_abline(slope = -0.1, intercept = 27, color = "blue",
              lwd = 1.5) +
  labs(x = "Miles", 
       y = "Price",
       title = "SSE = 1871.17") +
  theme_bw() +
  theme(text = element_text(size=25))

p2 <- AccordPrice %>%
  mutate(fitted = 20 - 0.1*Mileage) %>%
  ggplot(aes(x = Mileage, y = Price)) +
  geom_rect(aes(xmin = Mileage, 
                xmax = Mileage + Price - fitted,
                ymin = Price, ymax = fitted), 
            fill = "red", color = "red", alpha = 0.2) +
  geom_point(size=2.5) +
  geom_abline(slope = -0.1, intercept = 20, color = "blue",
              lwd = 1.5) +
  labs(x = "Miles", 
       y = "Price",
       title = "SSE = 287.49") +
  theme_bw() +
  theme(text = element_text(size=25))

p3 <- AccordPrice %>%
  mutate(fitted = 15 - 0.02*Mileage) %>%
  ggplot(aes(x = Mileage, y = Price)) +
  geom_rect(aes(xmin = Mileage, 
                xmax = Mileage + Price - fitted,
                ymin = Price, ymax = fitted), 
            fill = "red", color = "red", alpha = 0.2) +
  geom_point(size=2.5) +
  geom_abline(slope = -0.02, intercept = 15, color = "blue",
              lwd = 1.5) +
  labs(x = "Miles", 
       y = "Price",
       title = "SSE = 747.78") +
  theme_bw() +
  theme(text = element_text(size=25))

p1 + p2 + p3
```

.large[
Sum of squared residuals (SSE) = $\sum_{i=1}^n (y_i - \widehat{y}_i)^2 = \sum_{i=1}^n (y_i - \widehat{\beta}_0 - \widehat{\beta}_1 x_i)^2$

Choose $\widehat{\beta}_0$, $\widehat{\beta}_1$ to minimize SSE.
]

---

### Class activity, Part I

.large[
Spend a couple minutes trying out different potential values for the estimated slope and intercept. 

[https://sta112-f25.github.io/class_activities/ca_06.html](https://sta112-f25.github.io/class_activities/ca_06.html)
]

---

### Fitting linear regression in R

```{r echo=F, fig.width=6, fig.height=4, fig.align='center'}
AccordPrice %>%
  ggplot(aes(x = Mileage, y = Price)) +
  geom_point(size=2.5) +
  labs(x = "Miles", 
       y = "Price") +
  theme_bw() +
  theme(text = element_text(size=25))
```

.large[
```r
lm(Price ~ Mileage, data = AccordPrice)
```
]

.large[
* `lm`: function used for regression ("linear model")
* `Price`: response variable
* `Mileage`: predictor
]

---

### Fitting linear regression in R

.large[
```{r}
lm(Price ~ Mileage, data = AccordPrice)
```

* $\widehat{\beta}_0 = 20.81$
* $\widehat{\beta}_1 = -0.12$
]

---

### Class activity, Part II

.large[
[https://sta112-f25.github.io/class_activities/ca_06.html](https://sta112-f25.github.io/class_activities/ca_06.html)
]

---

### Calculating $\widehat{\beta}_0$ and $\widehat{\beta}_1$

.large[
.center[
$SSE = \sum_{i=1}^n (y_i - \widehat{\beta}_0 - \widehat{\beta}_1 x_i)^2$
]
]

.large[
Choose $\widehat{\beta}_0$, $\widehat{\beta}_1$ to minimize SSE

* Option 1: search for $\widehat{\beta}_0$, $\widehat{\beta}_1$ like in class activity Part I
]

---

### Calculating $\widehat{\beta}_0$ and $\widehat{\beta}_1$

.large[
.center[
$SSE = \sum_{i=1}^n (y_i - \widehat{\beta}_0 - \widehat{\beta}_1 x_i)^2$
]
]

.large[
Choose $\widehat{\beta}_0$, $\widehat{\beta}_1$ to minimize SSE

* Option 1: search for $\widehat{\beta}_0$, $\widehat{\beta}_1$ like in class activity Part I
]
.large[

    * not efficient
    * not what R does
]

---

### Calculating $\widehat{\beta}_0$ and $\widehat{\beta}_1$

.large[
.center[
$SSE = \sum_{i=1}^n (y_i - \widehat{\beta}_0 - \widehat{\beta}_1 x_i)^2$
]
]

.large[
Choose $\widehat{\beta}_0$, $\widehat{\beta}_1$ to minimize SSE

* Option 1: search for $\widehat{\beta}_0$, $\widehat{\beta}_1$ like in class activity Part I
]
.large[

    * not efficient
    * not what R does
]

.large[
* Option 2: closed-form equation?
]

---

### Equations for $\widehat{\beta}_0$ and $\widehat{\beta}_1$

.large[
.center[
$\widehat{\beta}_1 = \dfrac{\sum_{i=1}^n (x_i - \overline{x})(y_i - \overline{y})}{\sum_{i=1}^n (x_i - \overline{x})^2}$

$\widehat{\beta}_0 = \overline{y} - \widehat{\beta}_1 \overline{x}$
]
]

.large[
where 

.center[
$\overline{x} = \frac{1}{n} \sum_{i=1}^n x_i$

$\overline{y} = \frac{1}{n} \sum_{i=1}^n y_i$
]
]

---

### Making predictions

.large[
Goal: predict a house's sale price using either its year built or its size
]

```{r, echo=F, message=F, warning=F, fig.width=10, fig.height=4}
library(Stat2Data)
data("GrinnellHouses")
library(gridExtra)

p1 <- GrinnellHouses |>
  filter(YearBuilt > 1900) |>
  ggplot(aes(x = YearBuilt, y = SalePrice)) +
  geom_point() +
  geom_smooth(se=F, method = "lm") +
  theme_bw() +
  theme(text = element_text(size = 20)) +
  labs(x = "Year built",
       y = "Sale price ($)")

p2 <- GrinnellHouses |>
  filter(YearBuilt > 1900) |>
  ggplot(aes(x = SquareFeet, y = SalePrice)) +
  geom_point() +
  geom_smooth(se=F, method = "lm") +
  theme_bw() +
  theme(text = element_text(size = 20)) +
  labs(x = "House size (square feet)",
       y = "Sale price ($)")

grid.arrange(p1, p2, ncol = 2)
```

.large[
.question[
Which of the two variables would you prefer to use?
]
]

---

### Making predictions

.large[
.center[
$SSE = \sum_{i=1}^n (y_i - \widehat{y}_i)^2$
]
]

```{r, echo=F, message=F, warning=F, fig.width=10, fig.height=4}
library(Stat2Data)
data("GrinnellHouses")
library(gridExtra)

p1 <- GrinnellHouses |>
  filter(YearBuilt > 1900) |>
  ggplot(aes(x = YearBuilt, y = SalePrice)) +
  geom_point() +
  geom_smooth(se=F, method = "lm") +
  theme_bw() +
  theme(text = element_text(size = 20)) +
  labs(x = "Year built",
       y = "Sale price ($)")

p2 <- GrinnellHouses |>
  filter(YearBuilt > 1900) |>
  ggplot(aes(x = SquareFeet, y = SalePrice)) +
  geom_point() +
  geom_smooth(se=F, method = "lm") +
  theme_bw() +
  theme(text = element_text(size = 20)) +
  labs(x = "House size (square feet)",
       y = "Sale price ($)")

grid.arrange(p1, p2, ncol = 2)
```

.pull-left[
.large[
SSE: $3.41 \times 10^{12}$
]
]

.pull-right[
.large[
SSE: $1.74 \times 10^{12}$
]
]

.large[
SSE measures how close, overall, our predictions are to the observed response!
]

---

### Making predictions

.large[
.center[
$SSE = \sum_{i=1}^n (y_i - \widehat{y}_i)^2$
]
]

```{r, echo=F, message=F, warning=F, fig.width=10, fig.height=4}
library(Stat2Data)
data("GrinnellHouses")
library(gridExtra)

p1 <- GrinnellHouses |>
  filter(YearBuilt > 1900) |>
  ggplot(aes(x = YearBuilt, y = SalePrice)) +
  geom_point() +
  geom_smooth(se=F, method = "lm") +
  theme_bw() +
  theme(text = element_text(size = 20)) +
  labs(x = "Year built",
       y = "Sale price ($)")

p2 <- GrinnellHouses |>
  filter(YearBuilt > 1900) |>
  ggplot(aes(x = SquareFeet, y = SalePrice)) +
  geom_point() +
  geom_smooth(se=F, method = "lm") +
  theme_bw() +
  theme(text = element_text(size = 20)) +
  labs(x = "House size (square feet)",
       y = "Sale price ($)")

grid.arrange(p1, p2, ncol = 2)
```

.pull-left[
.large[
SSE: $3.41 \times 10^{12}$
]
]

.pull-right[
.large[
SSE: $1.74 \times 10^{12}$
]
]

.large[
.question[
Why are these numbers so *large*?
]
]

---

### Making predictions

.large[
.center[
$SSE = \sum_{i=1}^n (y_i - \widehat{y}_i)^2$
]
]

.large[
* Depends on the scale of the response variable
* Depends on the *number* of observations
]

