---
title: "Class Activity: categorical explanatory variables"
output: 
  tufte::tufte_html:
    css: "lab.css"
    tufte_variant: "envisioned"
    highlight: pygments
link-citations: yes
---

In this activity, you will practice fitting a regression model with a quantitative response and a single categorical predictor.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
drosophila <- read_csv("https://sta112-s22.github.io/class_activities/drosophila.csv")
diet_lm <- lm(WingLength ~ Diet, data = drosophila)
```

## Setup

Researchers are interested in investigating the relationship between diet and wing length for fruit flies (*Drosophila melanogaster*). In an experiment, they randomly assign infant flies to one of three diet groups: Standard, Rich, and Poor. They then measure each fly's wing length when they reach maturity.

You can load the data as follows:

```r
library(tidyverse)
drosophila <- read_csv("https://sta112-f25.github.io/class_activities/drosophila.csv")
```

Download the activity template: [activity template](https://sta112-f25.github.io/class_activities/ca_15_template.qmd)

### Modeling

We can visualize the relationship between diet and wing length with boxplots:

```{r echo=F, fig.align='center'}
drosophila %>%
  ggplot(aes(x = Diet, y = WingLength)) +
  geom_boxplot() +
  theme_bw() +
  theme(text = element_text(size = 20))
``` 

We will model this relationship as

$\text{WingLength} = \beta_0 + \beta_1 \text{IsRich} + \beta_2 \text{IsStandard} + \varepsilon$,

where $\text{IsRich}$ and $\text{IsStandard}$ are indicator variables for Diet:

```{marginfigure}
There are lots of different notation for indicator variables. For example, R would call these indicators `Dietrich` and `Dietstandard`. Other notation includes $I_{\text{rich}}$ and $I_{\text{standard}}$.
```

$\text{IsRich} = \begin{cases} 0 & \text{Diet } \neq \text{rich} \\ 1 & \text{Diet } = \text{rich} \end{cases}$

$\text{IsStandard} = \begin{cases} 0 & \text{Diet } \neq \text{standard} \\ 1 & \text{Diet } = \text{standard} \end{cases}$


## Questions

1. Use the `lm` function to fit the regression model with Diet as the predictor and WingLength as the response.

<SCRIPT>
function ShowAndHide(divid) {
    var x = document.getElementById(divid);
    if (x.style.display == 'none') {
        x.style.display = 'block';
    } else {
        x.style.display = 'none';
    }
}
</SCRIPT>

<BUTTON ONCLICK="ShowAndHide('q1_solution')">Solution</BUTTON>

<DIV ID="q1_solution" STYLE="display:none">
```r
diet_lm <- lm(WingLength ~ Diet, data = drosophila)
```
</DIV>

2. What is the estimated average WingLength for drosophila in each diet type?

<BUTTON ONCLICK="ShowAndHide('q2_solution')">Solution</BUTTON>

<DIV ID="q2_solution" STYLE="display:none">
```{r}
summary(diet_lm)
```

| Diet | $\widehat{\text{WingLength}}$ |
| --- | --- |
| Poor | 1.88 |
| Rich | 1.88 + 0.16 = 2.04 |
| Standard | 1.88 + 0.12 = 2.00 |

</DIV>

3. How many degrees of freedom does the fitted model from question 1 have?

<BUTTON ONCLICK="ShowAndHide('q4_solution')">Solution</BUTTON>

<DIV ID="q4_solution" STYLE="display:none">
The fitted model has 1369 = 1372 - 3 degrees of freedom (remember that we lose one degree of freedom for each parameter we estimate. We estimated $\beta_0, \beta_1, \beta_2$.)

