---
title: "Exam 2 review questions"
format: html
---

Below are questions to help you study for Exam 2. These are some examples of the kinds of questions I might ask.

* This is not a practice exam.
* The questions cover most, but not all, possible material for the exam.
* The distribution of questions here is not necessarily reflective of the distribution of questions on the actual exam.


## Multiple choice questions

1. Which of the following is the correct formula for $R^2$?
    (a) $R^2 = \dfrac{\sum \limits_{i=1}^n (y_i - \widehat{y}_i)^2}{\sum \limits_{i=1}^n (y_i - \overline{y})^2}$
    (b) $R^2 = 1 - \dfrac{\sum \limits_{i=1}^n (y_i - \widehat{y}_i)^2}{\sum \limits_{i=1}^n (y_i - \overline{y})^2}$
    (c) $R^2 = \dfrac{\frac{1}{n-p}  \sum \limits_{i=1}^n (y_i - \widehat{y}_i)^2}{ \frac{1}{n-1} \sum \limits_{i=1}^n (y_i - \overline{y})^2}$
    (d) $R^2 = 1 - \dfrac{\frac{1}{n-p}  \sum \limits_{i=1}^n (y_i - \widehat{y}_i)^2}{ \frac{1}{n-1} \sum \limits_{i=1}^n (y_i - \overline{y})^2}$
    (e) $R^2 = \dfrac{\frac{1}{p-1}  \sum \limits_{i=1}^n (y_i - \widehat{y}_i)^2}{ \frac{1}{n-1} \sum \limits_{i=1}^n (y_i - \overline{y})^2}$
    (f) $R^2 = 1 - \dfrac{\frac{1}{p-1}  \sum \limits_{i=1}^n (y_i - \widehat{y}_i)^2}{ \frac{1}{n-p} \sum \limits_{i=1}^n (y_i - \overline{y})^2}$
    
2. Which of the following is the correct formula for $R_{adj}^2$?
    (a) $R_{adj}^2 = \dfrac{\sum \limits_{i=1}^n (y_i - \widehat{y}_i)^2}{\sum \limits_{i=1}^n (y_i - \overline{y})^2}$
    (b) $R_{adj}^2 = 1 - \dfrac{\sum \limits_{i=1}^n (y_i - \widehat{y}_i)^2}{\sum \limits_{i=1}^n (y_i - \overline{y})^2}$
    (c) $R_{adj}^2 = \dfrac{\frac{1}{n-p}  \sum \limits_{i=1}^n (y_i - \widehat{y}_i)^2}{ \frac{1}{n-1} \sum \limits_{i=1}^n (y_i - \overline{y})^2}$
    (d) $R_{adj}^2 = 1 - \dfrac{\frac{1}{n-p}  \sum \limits_{i=1}^n (y_i - \widehat{y}_i)^2}{ \frac{1}{n-1} \sum \limits_{i=1}^n (y_i - \overline{y})^2}$
    (e) $R_{adj}^2 = \dfrac{\frac{1}{p-1}  \sum \limits_{i=1}^n (y_i - \widehat{y}_i)^2}{ \frac{1}{n-1} \sum \limits_{i=1}^n (y_i - \overline{y})^2}$
    (f) $R_{adj}^2 = 1 - \dfrac{\frac{1}{p-1}  \sum \limits_{i=1}^n (y_i - \widehat{y}_i)^2}{ \frac{1}{n-p} \sum \limits_{i=1}^n (y_i - \overline{y})^2}$
    
## Short answer questions

3. Can $R^2$ decrease when additional variables are added to a model? Explain. What about $R^2_{adj}$?

4. How should we handle potential outliers with a studentized residual $< -3$ or $> 3$? 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)

set.seed(3)

x = rnorm(100, mean=50, sd=15)
crabs <- data.frame(Noise = rep(c("ocean", "ship"), each = 50),
                       Mass = x,
                       Oxygen = c(50 + 2*x[1:50] + rnorm(50, sd=15),
                                  300 - 2*x[51:100] + rnorm(50, sd=15)))
```

## Example data questions

How does stress impact animals' oxygen consumption? To investigate this question, biologists exposed 100 crabs to either 7.5 minutes of loud ship noise, or 7.5 minutes of normal ocean noise, and recorded their oxygen consumption. The data contain the following variables.

* `Mass`: crab mass (g)
* `Oxygen`: rate of oxygen consumption (micromoles per hour)
* `Noise`: type of noise (either `ocean` or `ship`)

The researchers are interested in the relationship between mass and oxygen consumption, and the impact of noise on this relationship. They make the following scatterplot:

```{r, echo=F, fig.align='center', fig.width=6, fig.height=4}
crabs %>%
  ggplot(aes(x = Mass, y = Oxygen, 
             color = Noise, shape = Noise)) +
  geom_point(size=2) +
  theme_bw() +
  theme(text = element_text(size=20))
```

They consider three different models for this relationship ($\text{IsShip}$ is an indicator variable which is 1 for ship noise, and 0 for ocean noise):

$(1) \hspace{1cm} \text{oxygen} = \beta_0 + \beta_1 \text{mass} + \varepsilon$

$(2) \hspace{1cm} \text{oxygen} = \beta_0 + \beta_1 \text{IsShip} + \beta_2 \text{mass} + \varepsilon$

$(3) \hspace{1cm} \text{oxygen} = \beta_0 + \beta_1 \text{IsShip} + \beta_2 \text{mass} + \beta_3 \text{IsShip} \cdot \text{mass} +  \varepsilon$

The estimated equations for these models are

$(1) \hspace{1cm} \widehat{\text{oxygen}} = 161.29 + 0.23 \ \text{mass}$

$(2) \hspace{1cm} \widehat{\text{oxygen}} = 143.72 + 53.68 \ \text{IsShip} + 0.05 \ \text{mass}$

$(3) \hspace{1cm} \widehat{\text{oxygen}} = 52.06 + 257.38 \ \text{IsShip} + 1.92 \ \text{mass} - 4.05 \ \text{IsShip} \cdot \text{mass}$


1. Based on the scatterplot, which of these three models do you think is most appropriate? Explain your reasoning.

2. For each of the *other* two models, draw a scatterplot that would encourage you to choose that model, instead of the model selected in Question 1.

3. Using the estimated equation for the model you chose in question 1, what is the predicted rate of oxygen consumption for a 40g crab exposed to ocean noise? What about for a 40g crab exposed to ship noise?

4. What is the estimated change in oxygen consumption associated with a 1g increase in Mass, for crabs exposed to ship noise?

5. The researchers want to test whether there is any relationship between ship noise and oxygen consumption, after accounting for mass. Write down a null and alternative hypothesis for this test, in terms of one or more parameters from the model.

6. What test can the researchers use to test their hypotheses? Explain your reasoning, and provide the degrees of freedom.

7. Below are the ANOVA tables for the three fitted models. Use these tables to calculate a test statistic for the hypotheses in Question 5.

```{r}
m1 <- lm(Oxygen ~ Mass, data = crabs)
anova(m1)
```

<br/>

<br/>

```{r}
m2 <- lm(Oxygen ~ Mass + Noise, data = crabs)
anova(m2)
```

<br/>

<br/>

```{r}
m3 <- lm(Oxygen ~ Mass*Noise, data = crabs)
anova(m3)
```

8. Below is a picture of the distribution used to calculate a p-value with the test statistic in the previous question. Is the observed test statistic *usual* or *unusual* under the null hypothesis?






